### 2022-12-11 Paper/Overleaf/old_tamplate/samplepaper.tex
Added LaTeX Files

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%
\begin{document}
%
\title{Spicy short title}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Tim Gutberlet \and Janik Sauerbier}
%
\authorrunning{T. Gutberlet \& J. Sauerbier}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{University of Mannheim, Mannheim, Germany\\
\email{\{tgutberl,jsauerbi\}@mail.uni-mannheim.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Retrieving information from knowledge graphs is a highly relevant problem in academia and the industry. This includes among other things the knowledge graph completion/link prediction problem and rule-based pattern recognition. We focus on the problem of finding all rules that imply a certain target fact given a knowledge graph and a set of previously learned rules over the knowledge graph. Here we specifically analyze how relational database technologies perform in solving this problem, including a comparison of different indexing, hashing and pre-processing methods. Our research might help to advance approaches for link prediction and can be used to better understand the dynamics of rules given large amounts of target facts.

\keywords{Knowledge graph completion \and Link prediction  \and Knowledge representation and reasoning \and Database design and models.}
\end{abstract}
%
%
%
\section{Introduction}
Retrieving information from knowledge graphs has been a much-discussed topic in research over the years. Knowledge graphs are used in different fields from biomedical applications [OpenBioLink, HETIONET, Bio2RDF] over supply chain risk analysis to semantic search applications. Understanding patterns and rules as well as predicting links based on these patterns are relevant problems in the context of knowledge graphs in academia and industry. There are already many effective tools available that solve these problems, either as black box knowledge graph embedding models or through learning interpretable symbolic rules. We want to contribute to the field by analyzing on how to effectively find rules that imply certain target facts, which have not been part of the knowledge graph yet. We are not concerned with learning those rules, but just with identifying rules that have been learned previously and explain a certain target fact. Here, we specifically explore relational database technologies.

Many techniques are used to approach the link prediction problem, including various embeddings and rule-based methods. Prior research indicates that rule-based approaches and embedding models can be used together to improve inference quality. This can be done by combining embedding models and constraints derived from rules as the objective function of an integer linear programming problem, or by employing other methods focused on combining the results of different approaches. Another method is the use of rules during the training of embedding models. There is a significant need for a time-efficient database architecture which is used to identify rules that implicate the predictions of the embedding models.

Beyond potentially facilitating faster link prediction in the future, our findings can already be helpful to build tools that help to understand the dynamics of rules given large amounts of potential facts for knowledge graphs. Our research might also be helpful to understand the limits of relational database technologies in the context of knowledge graphs and might help to facilitate the comparison between different database technologies. […]

To achieve those goals, we ran several experiments aimed to illustrate how different database structures impact the performance in terms of execution time using different indexing, hashing and pre-processing methods. Beyond that, we also analyzed how certain queries (aka. the target facts) and certain rule types (e.g., by length of the rule body) performed in comparison to each other.

\section{Related Work}
\begin{itemize}
  \item Work on combining embeddings and rule-based approaches. (e.g. Wang, Quan, Bin Wang, and Li Guo. "Knowledge base completion using embeddings and rules."\cite{ref_3} or Zhang, Wen, et al. "Iteratively learning embeddings and rules for knowledge graph reasoning."\cite{ref_4})
  \item Work specifically focused on building effective databases (e.g. Indexing, Hashing and preprocessing strategies)
\end{itemize}

\section{Preliminaries}
\subsubsection{Knowledge Graphs and Rules} 
A knowledge graph (KG) is a set of triples (subject, relation, object) which describe a directed graph. The edges are represented by the relations going from subject to object. The nodes are the set of entities which appear as subjects and objects of the triples. The individual triples are also called facts. Knowledge graphs can be used to describe a variety of domains and can store knowledge about them. Possible domains are for example found in the biomedical field or in the context of the Semantic Web. Knowledge graphs can be used to identify rules and patterns in those domains. This can be very helpful to for example analyze how different biomedical processes correlate with certain genes, diseases, or treatments. For our purposes, we are concerned about first-order logic Horn rules. These rules consist of a head and a body, where the head consists of one triple and the body consists of one or more triples. The subjects and objects of the triples can either be specific entities or variables. Here are two example rules. We use the notation relation(subject, object) for the triples and capitalize the variables.

\begin{equation}
citizenOf(X, germany) \leftarrow bornIn(X, mannheim) 
\end{equation}
\begin{equation}
livesIn(X, germany) \leftarrow marriedTo(X, A_1), bornIn(A_1, germany)
\end{equation}

The grounding of a rule is a set of triples that fulfil the rule. That means the knowledge graph contains triples with matching entities for all variables. The triple set {livesIn(tom, germany), marriedTo(tom, bob), bornIn(bob, germany)} would be a grounding for the second rule. Rules are not necessarily universally true, and there might be sets of triples that match the body, but there is no matching triple for the head. Based on the ratio of the number of groundings of a rule (matching head + body) and the number of matching bodies without a matching head, we can derive a confidence score for any given rule.

\subsubsection{Knowledge Graph Completion} 
As knowledge graphs are often based on real-world data, they are prone to be incomplete. Therefore, one very relevant problem in the context of knowledge graphs is link prediction/knowledge graph completion. The problem is concerned with identifying missing links (aka. missing facts) that might be true in the real world but are not yet part of the knowledge graph. One example would be the generation of friendship suggestions for users of social media sites based on known connections. A variety of black box and interpretable approaches have been used to tackle the problem. These include knowledge graph embedding models and rules-based approaches.

\section{Problem statement}
The problem we are trying to solve is identifying rules that imply certain target facts that are not part of the knowledge graph. There are various rule learners like AnyBURL available that can learn many rules even on big knowledge graphs in a short amount of time. That’s why we are not concerned with learning those rules based on a given knowledge graph. Instead, we want to identify all matching rules for a certain target triple based on a previously learned set of rules. This means that the target triple should match the head of the rule and a subset of triples from the knowledge graph should match the body of the rule. We want to specifically explore in how far we can use traditional relational database technology to solve the problem as efficiently as possible. To illustrate the problem, think of the following example. We are given a knowledge graph, a target triple and a set of rules previously learned over the knowledge graph.

\begin{itemize}
  \item \textbf{Knowledge Graph:}
  \item[] \(\{…, marriedTo(anna, peter), bornIn(peter, germany), …\}\)
  \item \textbf{Target Fact:}
  \item[] \(livesIn(anna, germany)\)
  \item \textbf{Rules:}
  \begin{enumerate}
      \item \(citizenOf(X, germany) \leftarrow bornIn(X, mannheim)\)
      \item \(livesIn(X, germany) \leftarrow marriedTo(X, A_1), bornIn(A_1, germany)\)
  \end{enumerate}
\end{itemize}

Our goal is to identify all rules that imply the target fact. In our example we can start off with excluding rule one from further investigation as the target fact needs to match the head of the rule. Rule two could potentially imply the target fact with \(X = anna\). To investigate whether rule two actually implies the target fact, we need to find matching facts for the rule body from the knowledge graph. The knowledge graph contains the facts \(marriedTo(anna, peter)\) and \(bornIn(peter, germany)\). Together with the target fact, this results in a grounding for the second rule with \(X = anna\) and \(A_1 = peter\). Therefor, the solution for the example problem is rule two.

\section{Implementation}

\subsubsection{Summary} 
Our tool was implemented using the open-source object-relational database system PostgreSQL. Before accepting any queries, we store the rules into a hash map to allow for easy pre-filtering of the rules. The given knowledge graph will be stored in tables of the relational database. After that, our implementation is ready to accept queries which each contain a single target fact. The result is the set of rules implying the target fact.

\subsubsection{Data Model of the Rules} 
Before processing any queries, our tool sorts the given set of rules into a hash map with five keys. The keys are based on the structure of the head of the rule.

\begin{enumerate}
  \item \textbf{Subject Bound:} The head has a fixed entity as the subject and a variable as the object.
  \item \textbf{Object Bound:} The head has a variable as the subject and a fixed entity as the object.
  \item \textbf{Both Bound:} The head has a fixed entity as the subject and the object.
  \item \textbf{Unbound and Unequal:} The head has a variable as the subject and another variable as the object.
  \item \textbf{Unbound and Equal:} The head has a variable as the subject and the same variable as the object.
\end{enumerate}

\noindent
Within the respective categories, the rules are again sorted into a hash map by all fixed components (relation + potentially subject and object). This allows for initial filtering of the rules in O(1).

\subsubsection{Data Model of the Knowledge Graph}
The knowledge graph consisting of \((subject, relation, object)\) triples is stored in a table with three columns for each element of the triple (column names: “sub” = subjects, “pre” = relations, “obj” = objects).
We tested three different indexing approaches to optimize the execution time of the queries:

\begin{enumerate}
  \item \textbf{Secondary Unclustered Index} on all three columns (sub, pre, obj)
  \item \textbf{Unique Primary Clustered Index} on all three columns (sub, pre, obj)
  \item \textbf{Secondary Unclustered Index} on the column combinations (pre), (pre, obj), (sub, pre) and (sub, pre, obj)
\end{enumerate}

Beyond that, we created an alternative approach using materialized views (aka. additional tables) for each relation in the knowledge graph. These views only contain two columns for the subject and the object. Instead of checking in the overall knowledge graph table if a rule applies, we now check if a specific triple exists in its respective materialized view. We tested this alternative approach once without indexes and once with a Unique Unclustered Secondary Index on all columns (subject + object).

\subsubsection{Queries}
The queries for our initial data model of the knowledge graph are built as follows. The basic idea is to check every possible rule that wasn't filtered out before. The SQL statement is an \(UNION ALL\) operation over all subqueries for the individual rules. The SQL statement structure of the subquery is implemented as one \(SELECT\) statement. \(N\) is the number of body triples of the checked rule and \(rule\_id\) is a unique identifier for the checked rule. \(kg\_table\) is the name of the table containing all triples of the knowledge graph. 
\newline

\noindent
\textit{(\textbf{SELECT} $rule\_id_1$ \textbf{FROM} kg\_table t0, kg\_table t1, …, kg\_table tN \textbf{WHERE}  t0.sub = c0 \textbf{AND} t0.obj = c1 \textbf{AND} t1.sub = c2 \textbf{AND} t1.obj = c3 \textbf{AND} t0.obj = t1.sub … \textbf{AND} tN.sub = cN-1 and tN.obj = cN \textbf{AND} tN-1.obj = tN.sub \textbf{LIMIT} 1) \textbf{UNION ALL} (\textbf{SELECT} $rule\_id_2$ …) …}
\newline

\noindent
For the first alternative setup using one materialized view per relation, the SQL statement is structured very similar. The only difference is lies in the fact that the table \(kg\_table\) is now split up by relation. The table \(kg\_table_{rX}\) contains all triples of the relevant relation at the respective position in the rule body.
\newline

\noindent
\textit{(\textbf{SELECT} $rule\_id_1$ \textbf{FROM} $kg\_table_{r1}$ t0, $kg\_table_{r2}$ t1, …, $kg\_table_{rN}$ tN \textbf{WHERE}  t0.sub = c0 \textbf{AND} t0.obj = c1 \textbf{AND} t1.sub = c2 \textbf{AND} t1.obj = c3 \textbf{AND} t0.obj = t1.sub … \textbf{AND} tN.sub = cN-1 and tN.obj = cN \textbf{AND} tN-1.obj = tN.sub \textbf{LIMIT} 1) \textbf{UNION ALL} (\textbf{SELECT} $rule\_id_2$ …) …}

\section{Experiments}
%For our experiments, we used YAGO3-10 as a benchmark dataset. The dataset was divided into a training dataset (1079040 triples), test dataset (5000 triples) and query dataset (5000 triples). We used the AnyBURL implementation for the bottom up rule-based learning approach learning rules for 50 seconds, only generating rules that do not have empty bodies (107301 Rules). To have an independent set of queries for which we want to test if they fire, we used the query dataset.

Our approach is written in Java and requires a PostgreSQL JDBC Driver installed. The source code and datasets, used in the experiments, can be found at https://github.com/timgutberlet/Online-Rule-Search-Database. We conducted all experiments
on a Fujitsu Esprimo P957 (construction year 2017) with 32 GB RAM, 512 GB SSD, an Intel i7-7700 @ 3.6 GHz CPU and Ubuntu 22.04.1 LTS as an operating system.

%AnyBURL offers a similar feature for link prediction, which we will use to compare our optimisation features with. For the Comparison with AnyBURL given the limitation of AnyBURL to only find the following rule types:
%\begin{equation}
%h(X,Y) \leftarrow r(X,Y)
%\end{equation}
%\begin{equation}
%h(X,Y) \leftarrow r_1(X,Z) \wedge r_2(Z,Y)
%\end{equation}
%\begin{equation}
%h(X,e_1) \leftarrow r(X,e_2)
%\end{equation}
%Using the given rule set and filtering it by these easy rule types, this leaves 19271 Rules from the original 106480 rules. 
%As we have identified, our approach of [...] is most efficient for rule sets that only contain easy rule types, like in the AnyBURL example. Testing the approach of AnyBURL it takes on average 11 ms per query to find all rules, if there are any, that determine if the query fires. For our approach, it takes on average 3 ms. To find out the performance influences of different optimisation features, we deactivated each feature on an isolated level and looked at how this worsened performance. [...]

%However, to show the relevance of our approach, we also tested the performance of our optimisation strategies, without limiting the rule types in any sense.

%Conclusions / Assumptions:
%Yago3 - 10 has relatively low amount of 37 different relations. Working with datasets with a higher count of relations could possibly improve the approach of using materialized views for every relation, as the knowledgegraph would be split even more fine-grained.

\section{Conclusions}

%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%

\subsubsection{Acknowledgements} ...

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_1_adversarial_explanations_betz}
Betz, Patrick, Christian Meilicke, and Heiner Stuckenschmidt. "Adversarial explanations for knowledge graph embeddings." International Joint Conferences on Artificial Intelligence, 2022.
\bibitem{ref_2_abductive_reasoning}
Mayer, Marta Cialdea, and Fiora Pirri. "First order abduction via tableau and sequent calculi." Logic Journal of the IGPL 1.1 (1993): 99-117.
\bibitem{ref_3}
Wang, Quan, Bin Wang, and Li Guo. "Knowledge base completion using embeddings and rules." Twenty-fourth international joint conference on artificial intelligence. 2015.
\bibitem{ref_4}
Zhang, Wen, et al. "Iteratively learning embeddings and rules for knowledge graph reasoning." The World Wide Web Conference. 2019.

\end{thebibliography}
\end{document}

### 2022-12-11 Code/sql-implementations/src/app/Main.java
Added Native Implementation without pre filtering

@@ -42,7 +42,12 @@ public class Main {
     } else if(Config.getStringValue("TESTRULES_METHOD").equals("optimizedQuantileAnalysis") && Config.getStringValue("REFILL_TABLES").equals("YES")){
       randomRules.learnRules();
     }
-    randomRules.startQuery();
+    if(!Config.getStringValue("NATIVE").equals("YES")){
+      randomRules.startQuery();
+    }else {
+      System.out.println("NATIVE!!!");
+      randomRules.startQueryNative();
+    }
 
     //CreateDB.createKnowledgeGraphDB();
     // DBFuncs.readAllData();

### 2022-12-11 Code/sql-implementations/src/database/DBFuncs.java

@@ -720,7 +720,7 @@ public class DBFuncs {
         ArrayList<Integer> foundRules = new ArrayList<>();
         boolean first, first2;
         Triple r1, r2;
-
+        Boolean optimizeActivated = "optimizedQuantileAnalysis".equals(Config.getStringValue("TESTRULES_METHOD"));
         ResultSet rs;
         try {
             Statement stmt = con.createStatement();
@@ -728,6 +728,7 @@ public class DBFuncs {
             sql = new StringBuffer();
             StringBuffer sub, obj;
             StringBuffer select, where;
+            int iZahl;
             int help;
             boolean firstR = true;
             for (Rule rule : filteredRules) {
@@ -742,173 +743,187 @@ public class DBFuncs {
                 } else {
                     sql.append(" UNION ALL (");
                 }
-
-                sqlEnd = new StringBuffer(" LIMIT 1) ");
-                select = new StringBuffer("SELECT " + rule.getId() + " FROM ");
-                where = new StringBuffer(" WHERE");
-                help = 0;
-                for (Triple triple : rule.getBody()) {
-                    switch (Config.getStringValue("PREDICATE_VIEW")) {
-                        case "viewsForPredicate":
-                            if (triple.getSubject() >= 0) {
-                                if (first) {
-                                    select.append("r" + triple.getPredicate() + " k" + help);
-                                    first = false;
-                                } else {
-                                    select.append(", " + "r" + triple.getPredicate() + " k" + help);
-                                }
-                            } else if (triple.getObject() >= 0) {
-                                if (first) {
-                                    select.append("l" + triple.getPredicate() + " k" + help);
-                                    first = false;
-                                } else {
-                                    select.append(", " + "l" + triple.getPredicate() + " k" + help);
-                                }
-                            } else if (triple.getSubject() == rule.getHead().getSubject() || triple.getSubject() == rule.getHead().getObject()) {
-                                if (first) {
-                                    select.append("l" + triple.getPredicate() + " k" + help);
-                                    first = false;
-                                } else {
-                                    select.append(", " + "l" + triple.getPredicate() + " k" + help);
-                                }
-                            } else if (triple.getObject() == rule.getHead().getSubject() || triple.getObject() == rule.getHead().getObject()) {
-                                if (first) {
-                                    select.append("r" + triple.getPredicate() + " k" + help);
-                                    first = false;
-                                } else {
-                                    select.append(", " + "r" + triple.getPredicate() + " k" + help);
-                                }
[…]

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java

@@ -314,6 +314,63 @@ public class RandomRules {
         return string;
     }
 
+    public ArrayList<Integer> searchByTripleNative(Triple triple){
+        int count = 0;
+        ArrayList<Integer> resultList =  new ArrayList<>();
+        ArrayList<Rule> filteredRules = new ArrayList<>();
+        for (Map.Entry<Key2Int, ArrayList<Rule>> entry : objBound.entrySet()) {
+            for(Rule r: entry.getValue()){
+                if(triple.getPredicate() == r.getHead().getPredicate() && triple.getObject() == r.getHead().getObject()){
+                    filteredRules.add(r);
+                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
+                    //System.out.println(count++);
+                }
+                filteredRules.clear();
+            }
+        }
+        for (Map.Entry<Key2Int, ArrayList<Rule>> entry : subBound.entrySet()) {
+            for(Rule r: entry.getValue()){
+                if(triple.getPredicate() == r.getHead().getPredicate() && triple.getSubject() == r.getHead().getSubject()){
+                    filteredRules.add(r);
+                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
+                    //System.out.println(count++);
+                }
+                filteredRules.clear();
+            }
+        }
+        for (Map.Entry<Key3Int, ArrayList<Rule>> entry : bothBound.entrySet()) {
+            for(Rule r: entry.getValue()){
+                if(triple.getPredicate() == r.getHead().getPredicate() && triple.getObject() == r.getHead().getObject() && triple.getSubject() == r.getHead().getSubject()){
+                    filteredRules.add(r);
+                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
+                    //System.out.println(count++);
+                }
+                filteredRules.clear();
+            }
+        }
+        for (Map.Entry<Integer, ArrayList<Rule>> entry : noBoundEqual.entrySet()) {
+            for(Rule r: entry.getValue()){
+                if(triple.getSubject() == triple.getObject() && triple.getPredicate() == r.getHead().getPredicate()){
+                    filteredRules.add(r);
+                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
[…]

### 2022-12-11 Paper/Overleaf/paper.tex

% !TeX encoding = UTF-8
% !TeX program = pdflatex
% !BIB program = bibtex

%%% Um einen Artikel auf deutsch zu schreiben, genügt es die Klasse ohne
%%% Parameter zu laden.
\documentclass[english]{lni}
%%% To write an article in English, please use the option ``english'' in order
%%% to get the correct hyphenation patterns and terms.
%%% \documentclass[english]{class}
%%

\usepackage{booktabs}
\usepackage{pgfplots}

\pgfplotsset{width=12.5cm, height=4cm}

\begin{document}
%%% Mehrere Autoren werden durch \and voneinander getrennt.
%%% Die Fußnote enthält die Adresse sowie eine E-Mail-Adresse.
%%% Das optionale Argument (sofern angegeben) wird für die Kopfzeile verwendet.
\title[From Facts to Rules using Relational Databases]{From Facts to Rules using Relational Databases}
%%% \subtitle{From Facts to Rules using Relational Databases} % if needed
\author[Tim Gutberlet \and Janik Sauerbier]
{Tim Gutberlet\footnote{University of Mannheim, \email{tgutberl@mail.uni-mannheim.de}} \and
Janik Sauerbier\footnote{University of Mannheim, \email{jsauerbi@mail.uni-mannheim.de}}}
\startpage{1} % Beginn der Seitenzählung für diesen Beitrag / Start page
\editor{K.-U. Sattler et al.} % Names of Editors
\booktitle{Datenbanksysteme für Business, Technologie und Web (BTW 2023)} % Name of book title
\yearofpublication{2023}
%%%\lnidoi{18.18420/provided-by-editor-02} % if known
\maketitle

\begin{abstract}
In this paper, we focus on the problem of finding all rules that imply a certain target fact given a knowledge graph and a set of previously learned rules over the knowledge graph. We specifically analyze how relational database technology performs in solving this problem using indexing, filtering and pre-processing methods. Our research might help to advance approaches for link prediction and can be used to efficiently explain the dynamics of rules given large amounts of target facts or large knowledge graphs.

Test
\end{abstract}
\begin{keywords}
Knowledge graph completion \and Link prediction \and Knowledge representation and reasoning \and Relational databases.
\end{keywords}
%%% Beginn des Artikeltexts
\section{Introduction}
Knowledge graphs (KGs) are used in different fields from biomedical applications \cite{OpenBioLink} over applications in the context of social networks \cite{SocialNetworks} to semantic search applications\cite{SemanticSearch}. We want to contribute by analyzing how to effectively check which rules of a KG imply certain target facts using relational database technology. For the rule learning, we use AnyBURL, a fast bottom up rule learner for KGs.\cite{AnyBURL19}

Our research might be used as a primitive for the link prediction problem. Prior research indicates that rule-based approaches and embedding models can be used in combination to improve inference quality.\cite{RuleEmbeddingCombination1}\cite{RuleEmbeddingCombination2} There is a significant need for a time-efficient solution that is used to check which rules implicate certain facts. Beyond potentially facilitating faster link prediction, our findings can already be helpful to explain the dynamics of rules given large amounts of potential facts for KGs. It might also particularly help in case of large knowledge graphs and memory constraints.

To achieve those goals, we ran several experiments aimed to illustrate how different database setups impact the performance in terms of execution time using filtering, indexing and pre-processing methods. Beyond that, we also performed a quantile performance analysis on the queries and rules.

\section{Preliminaries} 
A knowledge graph is a set of triples (subject, relation, object) that describe a directed graph. The edges are represented by the relations going from subject to object. The nodes are the set of entities which appear as subjects and objects of the triples. The individual triples are also called facts. KGs can be used to describe various domains and to identify rules and patterns in those domains. For our purposes, we are concerned about first-order logic Horn rules. These rules consist of a head and a body, where the head consists of one triple and the body consists of one or more triples. The subjects and objects of the triples can either be specific entities or variables. Here are two example rules. We use the notation relation(subject, object) for the triples and capitalize the variables.

\begin{equation}
citizenOf(X, germany) \leftarrow bornIn(X, mannheim) 
\end{equation}
\begin{equation}
livesIn(X, germany) \leftarrow marriedTo(X, A_1), bornIn(A_1, germany)
\end{equation}

The grounding of a rule is a set of triples of the KG that fulfil the rule, with matching entities for all variables. The triple set \textit{\{livesIn(tom, germany), marriedTo(tom, bob), bornIn(bob, germany)\}} would be a grounding for the second rule.

\section{Problem Statement}
The problem we are trying to solve is identifying rules that imply certain target facts which are not part of the KG based on the KG and a previously learned set of rules. This means we are looking for a grounding of the rule with the target triple matching the head of the rule and a subset of triples from the KG matching the body of the rule. To illustrate the problem, think of the following example.

\begin{itemize}
  \item \textbf{Knowledge Graph:}
  \item[] \(\{…, marriedTo(anna, peter), bornIn(peter, germany), …\}\)
  \item \textbf{Target Fact:} \(livesIn(anna, germany)\)
  \item \textbf{Learned Rules:}
  \begin{enumerate}
      \item \(citizenOf(X, germany) \leftarrow bornIn(X, mannheim)\)
      \item \(livesIn(X, germany) \leftarrow marriedTo(X, A_1), bornIn(A_1, germany)\)
  \end{enumerate}
\end{itemize}

In our example we can start off with excluding rule one from further investigation as the target fact needs to match the head of the rule. Rule two has a matching head with \(X = anna\). To investigate whether rule two actually implies the target fact, we need to find matching facts for the rule body from the KG. The KG contains the facts \(marriedTo(anna, peter)\) and \(bornIn(peter, germany)\). Together with the target fact, this results in a grounding for the second rule with \(X = anna\) and \(A_1 = peter\). Therefore, rule two is part of the solution.

\section{Implementation}

Our solution was implemented using the open-source object-relational database system PostgreSQL. The basic idea behind our implementation is creating SQL queries for each target fact. The given KG will be stored in one table of the relational database (columns: “sub” for subjects, “pre” for relations and “obj” for objects). The SQL statement consists of \textit{UNION ALL} operations over all sub-queries for the individual rules. The rules are saved in a list and in a loop over all rules, the sub-queries for rules with matching heads are created. To improve this basic idea, we employed several optimizations listed below. This would be the query for the example given in the problem statement:

\textit{(\textbf{SELECT} rule\_1 \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} k0.sub = anna \textbf{AND} k0.obj = k1.sub \textbf{AND} k1.obj = germany \textbf{LIMIT} 1)  \textbf{UNION ALL} (...)}

\subsection{Hash Maps for Rule Pre-filtering} 
To speed up the pre-filtering of the rules, we sorted the given set of rules into a hash map with five keys based on the rule head type. Three of the head types have fixed entities, either as subject and/or as object [e.g., \textit{r1(e1, X)}, \textit{r1(X, e1)} or \textit{r1(e1, e2)}]. Two of the head types have a variable as subject and as object, either the same variable or different variables [e.g., \textit{r1(X, X)}, \textit{r1(X, Y)}]. Within the respective categories, the rules are again sorted into hash maps based on the relation and the fixed entities. This allows for initial filtering of the rules in O(1).

\subsection{Tables for each Relation}
We created an alternative approach to having one big table for all facts of the KG. This approach uses one table for each relation in the KG therefore decreasing the individual table size. They only contain two columns for the subject and the object. This solution improved the overall performance by reducing the size of the B-trees of the indexes.

\subsection{Indexing}
To speed up the joining of the tables, we duplicated the knowledge graph tables and employed a unique clustered index, once with (subject, object) as key and once with (object, subject) as key. Within the SQL statement, the tables with the order (subject, object) are used in case of a fixed subject and the tables with the order (object, subject) are used in the case of a fixed object or no fixed subject and object. This ensures that an index is used as often as possible. The index speeds up the search and enables the direct access to the table contents through the B-trees.

\subsection{Pre-processing of expensive Rules}
As we will illustrate in our experiments, certain rules are way more expensive than others. To address that, we integrated a simple pre-processing method. We gather an independent set of n facts and used them as target facts. The queries were amended with the \textit{EXPLAIN ANALYZE} command to get the execution time for the individual rule sub-queries. Afterwards, we ranked all rules by the sum of the execution time over all queries. The top x\% of rules were then pre-processed, by calculating all potential rule heads which could form a grounding together with a set of triples from the KG as rule body. We excluded rules with one body triple, as pre-processing didn’t improve the performance. We also excluded rules with more than two body triples, as those rules incur a pre-processing time two orders of magnitude higher than rules with two body triples, while they only make up a small fraction of the most expensive and frequent rules.

\section{Experiments}
The goal of our experiments was to benchmark the performance of our solution and to measure the effects of different optimizations. We also analyzed the execution time for individual queries and rules. Beyond that, we tested our solution on different datasets and rulesets. Our solution is written in Java using the PostgreSQL JDBC driver. The source code and datasets, used in the experiments, can be found at https://github.com/timgutberlet/From-Facts-to-Rules-using-Relational-Databases. We conducted all experiments on a Fujitsu Esprimo P957 (construction year 2017) with 32 GB RAM, 512 GB SSD, an Intel i7-7700 @ 3.6 GHz CPU and Ubuntu 22.04.1 LTS as an operating system.

\subsection{Ablation Study - Optimizations}

\begin{table}
\centering
\begin{tabular}{lll}
\toprule
Experiment & Avg. Execution Time \\
\midrule
All optimizations enabled & 10 ms\\
\midrule
Hash Maps for Rule Pre-filtering disabled & XX ms\\
Tables for each relation disabled & 140 ms\\
Indexing disabled & 1680 ms\\
Pre-processing of expensive rules disabled & 11 ms\\

\bottomrule
\end{tabular}
\caption{Ablation Study of Optimizations using YAGO3-10}
\label{tab:ablation}
\end{table}

For these experiments, we used YAGO3-10 as the benchmark dataset.\cite{YAGO3} The dataset consists of a training set (1079040 triples), test set and validation set (5000 triples each). We used AnyBURL to learn rules for 50 seconds, only generating rules that do not have empty bodies (107301 Rules). We used the training set as the KG and the validation set as the target triples. For the \textit{Pre-processing of expensive Rules} optimization, we use the test set (n = 5000) to create the rule ranking and pre-process the top 1\% rules. The results in \Cref{tab:ablation} are based on executing all 5000 queries.

The \textit{Tables for each Relation} optimization specifically reduced the execution time for long rules. For rules with one body triple, the execution time was reduced by 12\%, for two body triples by 25\% and for three body triples by 74\%. This is caused by the increased use of the indexes during execution.

\subsection{Quantile Performance Analysis}

\begin{tikzpicture}
\begin{axis}[
        title={Fig. 1: Query quantile analysis (All optimizations enabled)},
	x tick label style={
		/pgf/number format/1000 sep=},
	ylabel=Execution Time (ET) in ms,
        xlabel=Percentile,
	enlargelimits=0.05,
	legend style={at={(0.5,-0.1)},
	anchor=north,legend columns=-1},
	ybar interval=0.5,
]
\addplot 
	coordinates {(0, 0.984156) (5, 4.113395)
            (10, 4.242852) (15, 4.38007)
            (20, 4.5406935) (25, 4.6855745)
            (30, 4.860181) (35, 5.1234675)
            (40, 5.626696) (45, 8.143013)
            (50, 8.648211) (55, 8.905306)
            (60, 9.140784) (65, 9.355278)
            (70, 9.636537) (75, 9.959422)
            (80, 10.475147) (85, 17.016464)
            (90, 30.580482) (95, 32.047903) 
            (100, 0)};
\end{axis}
\label{Test}
\end{tikzpicture}

\begin{tikzpicture}
\begin{axis}[
        title={Fig. 2: Rule quantile analysis (All optimizations enabled)},
	x tick label style={
		/pgf/number format/1000 sep=},
	ylabel=1/Frequency * Avg. ET in ms,
        xlabel=Percentile,
	enlargelimits=0.05,
	legend style={at={(0.5,-0.1)},
	anchor=north,legend columns=-1},
	ybar interval=0.5,
]
\addplot 
	coordinates {(0, 0.021) (5, 0.025)
            (10, 0.026) (15, 0.027)
            (20, 0.027) (25, 0.028)
            (30, 0.030) (35, 0.031)
            (40, 0.032) (45, 0.036)
            (50, 0.042) (55, 0.052)
            (60, 0.057) (65, 0.064)
            (70, 0.08) (75, 0.1)
            (80, 0.152) (85, 1.64)
            (90, 8.26) (95, 12.90) 
            (100, 0) };
\end{axis}
\label{Test}
\end{tikzpicture}

The analysis indicates that only few rules and few queries incur a major share of the cost. The \textit{Pre-processing of expensive Rules} optimization helped to reduce the cost for very expensive rules. The execution time for a specific rule is influenced by the length of the rule, as well as the target triple and the structure of the knowledge graph.

\subsection{Further Datasets and Rulesets}

\begin{table}
\centering
\begin{tabular}{llll}
\toprule
 & AnyBURL - 10s & AnyBURL - 50s & AnyBURL - 100s\\
\midrule
YAGO3-10 & XX ms & 10 ms & XX ms\\
WN18RR & XX ms & XX ms & XX ms\\
FB15k-237 & XX ms & XX ms & XX ms\\
\bottomrule
\end{tabular}
\caption{Average Execution Time for different Datasets and Rulesets}
\label{tab:further-datasets}
\end{table}

Our solution works for different datasets and rulesets learned by AnyBURL in the given time. The solution is particularly well suited for datasets with many relations, like the Freebase dataset, as it allows for small B-trees of the indexes.

%However, to show the relevance of our approach, we also tested the performance of our optimisation strategies, without limiting the rule types in any sense.

%Conclusions / Assumptions:
%Yago3 - 10 has relatively low amount of 37 different relations. Working with datasets with a higher count of relations could possibly improve the approach of using materialized views for every relation, as the knowledgegraph would be split even more fine-grained.

\section{Conclusions}

%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%

We demonstrated an effective solution for finding all rules that imply a certain target fact given a knowledge graph and a set of previously learned rules. Our experiments specifically demonstrated the effect of filtering, indexing and pre-processing methods. Potential next steps include a further analysis of our solution on various datasets, a comparison of different database technologies (particularly triplestores), exploring the use of multithreading and creating a dedicated solution only using the main memory.

\textbf{Acknowledgement.} This paper would not have been possible without the exceptional support of our supervisor, Prof. Dr. Rainer Gemulla.

%%% Angabe der .bib-Datei (ohne Endung) / State .bib file (for BibTeX usage)
\bibliography{references} %\printbibliography if you use biblatex/Biber
\end{document}

### 2022-12-11 Paper/Overleaf/paper_sample.tex

% !TeX encoding = UTF-8
% !TeX spellcheck = de_DE

%% Dies gibt Warnungen aus, sollten veraltete LaTeX-Befehle verwendet werden
\RequirePackage[l2tabu, orthodox]{nag}

\documentclass[utf8,biblatex]{lni}
\bibliography{references_sample}

%% Schöne Tabellen mittels \toprule, \midrule, \bottomrule
\usepackage{booktabs}

%% Zu Demonstrationszwecken
\usepackage[math]{blindtext}
\usepackage{mwe}

%% BibLaTeX-Sonderkonfiguration,
%% falls man schnell eine existierende Bibliographie wiederverwenden will, aber nicht die .bib-Datei händisch anpassen möchte.
%% Bitte \iffalse und \fi entfernen, dann ist diese Konfiguration aktiviert.

\iffalse
\AtEveryBibitem{%
  \ifentrytype{article}{%
  }{%
    \clearfield{doi}%
    \clearfield{issn}%
    \clearfield{url}%
    \clearfield{urldate}%
  }%
  \ifentrytype{inproceedings}{%
  }{%
    \clearfield{doi}%
    \clearfield{issn}%
    \clearfield{url}%
    \clearfield{urldate}%
  }%
}
\fi

\begin{document}
%%% Mehrere Autoren werden durch \and voneinander getrennt.
%%% Die Fußnote enthält die Adresse sowie eine E-Mail-Adresse.
%%% Das optionale Argument (sofern angegeben) wird für die Kopfzeile verwendet.
\title[Ein Kurztitel]{Ein sehr langer Titel über mehrere Zeilen mit sehr vielen Worten und noch mehr Buchstaben}
%%%\subtitle{Untertitel / Subtitle} % falls benötigt
\author[Vorname1 Nachname1 \and Vorname2 Nachname2]
{Vorname1 Nachname1\footnote{Universität, Abteilung, Straße, Postleitzahl Ort, Land \email{emailaddress@author1}} \and
 Vorname2 Nachname2\footnote{University, Department, Address, Country \email{emailaddress@author2}}}
\startpage{11} % Beginn der Seitenzählung für diesen Beitrag
\editor{Herausgeber et al.}    % Namen der Herausgeber
\booktitle{Name-der-Konferenz} % Name des Tagungsband; optional Kurztitel
\yearofpublication{2017}
%%%\lnidoi{18.18420/provided-by-editor-02} % Falls bekannt
\maketitle

\begin{abstract}
Die \LaTeX-Klasse \texttt{lni} setzt die Layout-Vorgaben für Beiträge in LNI Konferenzbänden um.
Dieses Dokument beschreibt ihre Verwendung und ist ein Beispiel für die entsprechende Darstellung.
Der Abstract ist ein kurzer Überblick über die Arbeit der zwischen 70 und 150 Wörtern lang sein und das Wichtigste enthalten sollte.
Die Formatierung erfolgt automatisch innerhalb des abstract-Bereichs.
\end{abstract}

\begin{keywords}
LNI Guidelines \and \LaTeX Vorlage
\end{keywords}

\section{Verwendung}
Die GI gibt unter \url{http://www.gi-ev.de/LNI} Vorgaben für die Formatierung von Dokumenten in der LNI Reihe.
Für \LaTeX-Dokumente werden diese durch die Dokumentenklasse \texttt{lni} realisiert.

Dieses Dokument basiert auf der offiziellen Dokumentation, simplifiziert und setzt grundlegendes LaTeX-Wissen voraus.
Es werden generische Platzhalter an die entsprechenden Stellen (wie beispielsweise die Authoren-Angaben) gesetzt und nicht weiter an anderer Stelle dokumentiert.

Dieses Template ist wie folgt gegliedert:
\Cref{sec:demos} zeigt Demonstrationen der LNI-Verlage.
\Cref{sec:lniconformance} zeigt die Einhaltung der Richtlinien durch einfachen Text.

\section{Demonstrationen}
\label{sec:demos}
Das Symbol für Potenzmengen ($\powerset$) wird korrekt angezeigt.
Es ist kein Weierstraß-p ($\wp$) mehr.

Spitze Klammen können direkt eingegeben werden: <test />

Hier eine kleine Demonstration von \href{https://www.ctan.org/pkg/microtype}{microtype}:
\blindtext

\section{Demonstration der Einhaltung der Richtlinien}
\label{sec:lniconformance}

\subsection{Literaturverzeichnis}
Der letzte Abschnitt zeigt ein beispielhaftes Literaturverzeichnis für Bücher mit einem Autor \cite{Ez10} und zwei AutorInnen \cite{AB00}, einem Beitrag in Proceedings mit drei AutorInnen \cite{ABC01}, einem Beitrag in einem LNI Band mit mehr als drei AutorInnen \cite{Az09}, zwei Bücher mit den jeweils selben vier AutorInnen im selben Erscheinungsjahr \cite{Wa14} und \cite{Wa14b}, ein Journal \cite{Gl06}, eine Website \cite{GI19} bzw.\ anderweitige Literatur ohne konkrete AutorInnenschaft \cite{XX14}.
Es wird biblatex verwendet, da es UTF8 sauber unterstützt und \href{https://github.com/gi-ev/LNI/issues/5}{im Gegensatz zu lni.bst} keine Fehler beim bibtexen auftreten.

Referenzen sollten nicht direkt als Subjekt eingebunden werden, sondern immer nur durch Authorenanganben:
Beispiel: \Citet{AB00} geben ein Beispiel, aber auch \citet{Az09}.
Hinweis: Großes C bei \texttt{Citet}, wenn es am Satzanfang steht. Dies ist analog zu \texttt{Cref}.

Formatierung und Abkürzungen werden für die Referenzen \texttt{book}, \texttt{inbook}, \texttt{proceedings}, \texttt{inproceedings}, \texttt{article}, \texttt{online} und \texttt{misc} automatisch vorgenommen.
Mögliche Felder für Referenzen können der Beispieldatei \texttt{lni-paper-example-de.bib} entnommen werden.
Andere Referenzen sowie Felder müssen allenfalls nachträglich angepasst werden.

\subsection{Abbildungen}
\Cref{fig:demo} zeigt eine Abbildung.

\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{example-image}
  \caption{Demographik}
  \label{fig:demo}
\end{figure}

\subsection{Tabellen}
\Cref{tab:demo} zeigt eine Tabelle.

\begin{table}
\centering
\begin{tabular}{lll}
\toprule
Überschriftsebenen & Beispiel & Schriftgröße und -art \\
\midrule
Titel (linksbündig) & Der Titel \ldots & 14 pt, Fett\\
Überschrift 1 & 1 Einleitung & 12 pt, Fett\\
Überschrift 2 & 2.1 Titel & 10 pt, Fett\\
\bottomrule
\end{tabular}
\caption{Die Überschriftsarten}
\label{tab:demo}
\end{table}

\subsection{Programmcode}
Die LNI-Formatvorlage verlangt die Einrückung von Listings vom linken Rand.
In der \texttt{lni}-Dokumentenklasse ist dies für die \texttt{verbatim}-Umgebung realisiert.

\begin{verbatim}
public class Hello {
    public static void main (String[] args) {
        System.out.println("Hello World!");
    }
}
\end{verbatim}

Alternativ kann auch die \texttt{lstlisting}-Umgebung verwendet werden.

\Cref{L1} zeigt uns ein Beispiel, das mit Hilfe der \texttt{lstlisting}-Umgebung realisiert ist.

\begin{lstlisting}[caption={Beschreibung}, label=L1, language=Java]
public class Hello {
    public static void main (String[] args) {
        System.out.println("Hello World!");
    }
}
\end{lstlisting}

\subsection{Formeln und Gleichungen}

Die korrekte Einrückung und Nummerierung für Formeln ist bei den Umgebungen
\texttt{equation} und \texttt{align} gewährleistet.

\begin{equation}
  1=4-3
\end{equation}
und
\begin{align}
  2&=7-5\\
  3&=2-1
\end{align}

% \bibliography{lni-paper-example-de.tex} ist hier nicht erlaubt: biblatex erwartet dies bei der Preambel
%% Starten Sie "biber paper", um eine Biliographie zu erzeugen.
\printbibliography

\end{document}

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java
Added Native Implementation without pre filtering 2

@@ -315,58 +315,61 @@ public class RandomRules {
     }
 
     public ArrayList<Integer> searchByTripleNative(Triple triple){
-        int count = 0;
-        ArrayList<Integer> resultList =  new ArrayList<>();
         ArrayList<Rule> filteredRules = new ArrayList<>();
         for (Map.Entry<Key2Int, ArrayList<Rule>> entry : objBound.entrySet()) {
             for(Rule r: entry.getValue()){
                 if(triple.getPredicate() == r.getHead().getPredicate() && triple.getObject() == r.getHead().getObject()){
                     filteredRules.add(r);
-                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
-                    //System.out.println(count++);
                 }
-                filteredRules.clear();
             }
         }
         for (Map.Entry<Key2Int, ArrayList<Rule>> entry : subBound.entrySet()) {
             for(Rule r: entry.getValue()){
                 if(triple.getPredicate() == r.getHead().getPredicate() && triple.getSubject() == r.getHead().getSubject()){
                     filteredRules.add(r);
-                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
-                    //System.out.println(count++);
                 }
-                filteredRules.clear();
             }
         }
         for (Map.Entry<Key3Int, ArrayList<Rule>> entry : bothBound.entrySet()) {
             for(Rule r: entry.getValue()){
                 if(triple.getPredicate() == r.getHead().getPredicate() && triple.getObject() == r.getHead().getObject() && triple.getSubject() == r.getHead().getSubject()){
                     filteredRules.add(r);
-                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
-                    //System.out.println(count++);
                 }
-                filteredRules.clear();
             }
         }
         for (Map.Entry<Integer, ArrayList<Rule>> entry : noBoundEqual.entrySet()) {
             for(Rule r: entry.getValue()){
                 if(triple.getSubject() == triple.getObject() && triple.getPredicate() == r.getHead().getPredicate()){
                     filteredRules.add(r);
-                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
-                    //System.out.println(count++);
                 }
-                filteredRules.clear();
             }
         }
[…]

### 2022-12-11 Code/sql-implementations/src/app/Main.java
Fox in Learning Behavior

@@ -39,7 +39,7 @@ public class Main {
     System.out.println("RandomRules Set");
     if(Config.getStringValue("TESTRULES_METHOD").equals("optimizedQuantileAnalysis") && Config.getStringValue("REFILL_TABLES").equals("YES")){
       randomRules.learnQuery(dbFill.getSampleTriples());
-    } else if(Config.getStringValue("TESTRULES_METHOD").equals("optimizedQuantileAnalysis") && Config.getStringValue("REFILL_TABLES").equals("YES")){
+    } else if(Config.getStringValue("TESTRULES_METHOD").equals("optimizedQuantileAnalysis") && Config.getStringValue("REFILL_TABLES").equals("NO")){
       randomRules.learnRules();
     }
     if(!Config.getStringValue("NATIVE").equals("YES")){

### 2022-12-11 Code/sql-implementations/src/utils/DbFill.java

@@ -17,7 +17,7 @@ public class DbFill {
   private HashMap<String, Integer> predicateIndex;
   private HashMap<String, Integer> objectIndex;
 
-  private ArrayList<Triple> sampleTriples =  new ArrayList<>();
+  private ArrayList<Triple> sampleTriples = new ArrayList<>();
 
   public HashMap<String, Integer> getObjectIndex() {
     return objectIndex;
@@ -31,14 +31,17 @@ public class DbFill {
     return subjectIndex;
   }
 
+
+
   public ArrayList<Triple> getSampleTriples() {
+    this.setSampleTriples();
     return sampleTriples;
   }
 
   /**
    * Fills the KnowledgeGraph Table with all rows given in the knowledgegraph text file
    */
-  public void fillKnowledgegraph(){
+  public void fillKnowledgegraph() {
     String file = Config.getStringValue("KNOWLEDGEGRAPH");
     //DBFuncs.deleteKG();
     BufferedReader reader;
@@ -50,13 +53,8 @@ public class DbFill {
       List<Triple> kgList = new ArrayList<>();
       String subject, predicate, object;
       Integer subH, predH, objH;
-      int count = 1;
-      for (String line; (line = reader.readLine()) != null;) {
-        System.out.println(count++);
-        // Process line
-        //System.out.println(line);
+      for (String line; (line = reader.readLine()) != null; ) {
         triple = line.split("\\s");
-        //System.out.println(count++ + " - triple: " + triple[0] + ", " + triple[1] +", "+ triple[2]);
         subject = triple[0];
         predicate = triple[1];
         object = triple[2];
@@ -65,9 +63,7 @@ public class DbFill {
         objH = objectIndex.get(object);
         t = new Triple(subH, predH, objH);
         kgList.add(t);
-        if(count % 10 == 0 && sampleTriples.size() < Config.getIntValue("QUANTIL_LEARN_TRIPLES") ){
-          sampleTriples.add(t);
-        }
+
       }
       DBFuncs.deleteKG();
       DBFuncs.insertKnowledgegraph(kgList);
@@ -78,13 +74,82 @@ public class DbFill {
     }
   }
 
+  public void setSampleTriples() {
+    String file = Config.getStringValue("QUANTIL_LEARN_TRIPLES");
+    //DBFuncs.deleteKG();
+    BufferedReader reader;
+    Triple t;
+    try {
+      // java.io.InputStream
+      reader = new BufferedReader(new FileReader(file));
[…]

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java

@@ -14,8 +14,7 @@ import models.*;
  * @author timgutberlet
  */
 public class RandomRules {
-    private HashMap<Integer, Integer> ruleHashMap = new HashMap<>();
-    private HashMap<Integer, RuleTime> ruleTimeHashMap =  new HashMap<>();
+    private HashMap<Integer, RuleTime> ruleTimeHashMap = new HashMap<>();
     private HashMap<String, Integer> subjectIndex;
     private HashMap<String, Integer> predicateIndex;
     private HashMap<String, Integer> objectIndex;
@@ -44,6 +43,8 @@ public class RandomRules {
     public HashMap<Integer, ArrayList<Rule>> getNoBoundEqual() {
         return noBoundEqual;
     }
+
+    private int ruleCount;
 
     public RandomRules(HashMap<String, Integer> subjectIndex,
                        HashMap<String, Integer> predicateIndex, HashMap<String, Integer> objectIndex) {
@@ -120,24 +121,12 @@ public class RandomRules {
                                     objectID);
                             body.add(tripleHelp);
                         }
-                        //tailList.add(new Triple(Integer.parseInt(bodySubject), Integer.parseInt(delE(bodyObject)), Integer.parseInt(bodyPredicate)));
                     }
                 }
-                //Uncomment this, if you also want to inlcude empty rules
-        /*else {
-          tripleHelp = new Triple(DBFuncs.getSubjectID(bodySubject), DBFuncs.getPredicateID(bodyPredicate),
-              DBFuncs.getObjectID(bodyObject));
-          body.add(tripleHelp);
-          System.out.println(tripleHelp);
-        }*/
                 help1 = first.split("\\(", 2)[1].split(",", 2);
                 if (help1.length > 1) {
                     headSubject = help1[0];
                     headObject = help1[1].substring(0, help1[1].length() - 1);
-                }
-                //System.out.println(headPredicate + " : " + headSubject + " " + headObject + " <= " + second);
-                if (headSubject.length() == 1) {
-
                 }
                 if (headSubject.length() == 1) {
                     subjectID = Variables.getID(headSubject);
@@ -173,9 +162,9 @@ public class RandomRules {
                 if (Config.getStringValue("FILTER_SIMPLE_RULES").equals("YES")) {
                     continuer = false;
 
-                    if ( rule.getBody().size() == 4) {
-                            continuer = true;
-                            System.out.println(c++ + " Type2: " + rule);
[…]

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java
Calc changes for Quantil Analysis

@@ -629,7 +629,7 @@ public class RandomRules {
         System.out.println("Average: " + ((double) average) + " ms");
         System.out.println("Min: " + ((double) min) + " ms");
         System.out.println("Max: " + ((double) max) + " ms");
-        //System.exit(0);
+        System.exit(0);
     }
 
     public static boolean isFloat(double value) {
@@ -708,34 +708,22 @@ public class RandomRules {
         ArrayList<RuleTime> helpList = new ArrayList<>();
         ArrayList<Rule> ruleList = new ArrayList<>();
         ArrayList<Double> timeList = new ArrayList<>();
-        for (Map.Entry<Integer, RuleTime> entry : ruleTimeHashMap.entrySet()) {
-            helpList.add(entry.getValue());
-            timeList.add(entry.getValue().avg());
-        }
-        Collections.sort(helpList);
         int i = 0;
-        for (RuleTime ruleTime : helpList) {
-            System.out.println(" ");
-            System.out.println(ruleTime.avg());
-            System.out.println(ruleTime.sum());
-            System.out.println(ruleTime.getCount());
-            System.out.println(ruleTime.getRule());
-            if (ruleTime.getRule().getBody().size() == 2) {
-                ruleList.add(ruleTime.getRule());
+        for (Map.Entry<Integer, RuleTime> entry : ruleTimeHashMap.entrySet()) {
+            if(entry.getValue().getRule().getBody().size() == 2) {
                 i++;
+                helpList.add(entry.getValue());
+                timeList.add(entry.getValue().avg());
             }
-            if (i == ruleCount) {
+            if(i == ruleCount){
                 break;
             }
         }
+        Collections.sort(helpList);
         quantilCalcSum(timeList);
         DBFuncs.viewsForQuantiles(ruleList);
-        System.out.println("RuleList");
         for (Rule rule : ruleList) {
             rule.setLearned();
-            System.out.println(rule);
-            System.out.println(rule.getBound());
-            System.out.println(rule.isLearned());
         }
         rulePreSave(ruleList);
     }

### 2022-12-11 Code/sql-implementations/src/database/DBFuncs.java
Update Files, Update Config, Changes for Quantil Analysis for Rules

@@ -50,7 +50,6 @@ public class DBFuncs {
             long count = 0;
             for (Triple triple : list) {
                 count++;
-                //System.out.println(triple.toString());
                 stmt.setInt(1, triple.getSubject());
                 stmt.setInt(2, triple.getPredicate());
                 stmt.setInt(3, triple.getObject());
@@ -69,202 +68,14 @@ public class DBFuncs {
             con.commit();
             stmt.close();
             System.out.println("Success");
-
             System.out.println("Data has been inserted");
         } catch (SQLException e) {
             e.printStackTrace();
         }
     }
 
-    /**
-     * Inserts the vocabulary for subjects
-     *
-     * @param map
-     */
-    public static void insertSubjects(HashMap<String, Integer> map) {
-        PreparedStatement stmt;
-        try {
-            String sql = "INSERT INTO subjects(id, txt) VALUES (?,?)";
-            stmt = con.prepareStatement(sql);
-            long count = 0;
-            for (String txt : map.keySet()) {
-                count++;
-                //System.out.println(triple.toString());
-                stmt.setInt(1, map.get(txt));
-                stmt.setString(2, txt);
-                stmt.addBatch();
-                if (count % 500 == 0 || count == map.size()) {
-                    stmt.executeBatch();
-                    stmt.clearBatch();
-                }
-            }
-            stmt.executeBatch();
-            con.commit();
-            stmt.close();
-            System.out.println("Data has been inserted");
-        } catch (SQLException e) {
-            e.printStackTrace();
-        }
-    }
 
-    /**
-     * Inserts a predicate with its key in the database
-     *
-     * @param map
-     */
-    public static void insertPredicates(HashMap<String, Integer> map) {
-        PreparedStatement stmt;
-        try {
-            String sql = "INSERT INTO predicates(id, txt) VALUES (?,?)";
-            stmt = con.prepareStatement(sql);
-            long count = 0;
-            for (String txt : map.keySet()) {
-                count++;
-                //System.out.println(triple.toString());
-                stmt.setInt(1, map.get(txt));
-                stmt.setString(2, txt);
-                stmt.addBatch();
-                if (count % 500 == 0 || count == map.size()) {
-                    stmt.executeBatch();
-                    stmt.clearBatch();
-                }
-            }
-            stmt.executeBatch();
[…]

### 2022-12-11 Code/sql-implementations/src/models/Triple.java

@@ -7,6 +7,7 @@ package models;
  */
 public class Triple {
   private Integer subject, predicate, object;
+  String subString, preString, objString;
 
   /**
    * Constructor for setting a Triple
@@ -14,10 +15,13 @@ public class Triple {
    * @param predicate Integer
    * @param object Integer
    */
-  public Triple(Integer subject, Integer predicate, Integer object){
+  public Triple(Integer subject, Integer predicate, Integer object, String subString, String preString, String objString){
     this.subject = subject;
     this.predicate = predicate;
     this.object = object;
+    this.subString = subString;
+    this.preString = preString;
+    this.objString = objString;
   }
 
   /**
@@ -50,9 +54,16 @@ public class Triple {
    */
   @Override
   public String toString() {
-    return "Triple{" +
+    return "Triple: " +
          subject +
         " " + predicate +
         " " + object;
+  }
+
+  public String toText() {
+    return "Triple: " +
+            subString +
+            " " + preString +
+            " " + objString;
   }
 }

### 2022-12-11 Code/sql-implementations/src/utils/DbFill.java

@@ -13,9 +13,9 @@ import java.util.List;
 import models.Triple;
 
 public class DbFill {
-  private HashMap<String, Integer> subjectIndex;
-  private HashMap<String, Integer> predicateIndex;
-  private HashMap<String, Integer> objectIndex;
+  private HashMap<String, Integer> subjectIndex =  new HashMap<>();
+  private HashMap<String, Integer> predicateIndex =  new HashMap<>();
+  private HashMap<String, Integer> objectIndex =  new HashMap<>();
 
   private ArrayList<Triple> sampleTriples = new ArrayList<>();
 
@@ -53,6 +53,7 @@ public class DbFill {
       List<Triple> kgList = new ArrayList<>();
       String subject, predicate, object;
       Integer subH, predH, objH;
+      int helpCount = 0;
       for (String line; (line = reader.readLine()) != null; ) {
         triple = line.split("\\s");
         subject = triple[0];
@@ -61,9 +62,8 @@ public class DbFill {
         subH = subjectIndex.get(subject);
         predH = predicateIndex.get(predicate);
         objH = objectIndex.get(object);
-        t = new Triple(subH, predH, objH);
+        t = new Triple(subH, predH, objH, subject, predicate, object);
         kgList.add(t);
-
       }
       DBFuncs.deleteKG();
       DBFuncs.insertKnowledgegraph(kgList);
@@ -95,7 +95,7 @@ public class DbFill {
         subH = subjectIndex.get(subject);
         predH = predicateIndex.get(predicate);
         objH = objectIndex.get(object);
-        t = new Triple(subH, predH, objH);
+        t = new Triple(subH, predH, objH, subject, predicate, object);
         System.out.println(t);
         sampleTriples.add(t);
       }
@@ -110,55 +110,32 @@ public class DbFill {
    * Fills the Index Database files
    */
   public void fillVocabulary() {
-    String test = Config.getStringValue("QUERYTRIPLES");
-    //DBFuncs.deleteKG();
+    String kg = Config.getStringValue("KNOWLEDGEGRAPH");
     BufferedReader reader;
     try {
       // java.io.InputStream
-      reader = new BufferedReader(new FileReader(test));
+      reader = new BufferedReader(new FileReader(kg));
       String[] triple;
-      subjectIndex = new HashMap<>();
-      predicateIndex = new HashMap<>();
-      objectIndex = new HashMap<>();
       String subject, predicate, object;
       Integer subC = 0, predC = 0, objC = 0;
-      Integer count = 0;
[…]

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java

@@ -14,6 +14,7 @@ import models.*;
  * @author timgutberlet
  */
 public class RandomRules {
+    private ArrayList<Triple> validationTriples = new ArrayList<>();
     private HashMap<Integer, RuleTime> ruleTimeHashMap = new HashMap<>();
     private HashMap<String, Integer> subjectIndex;
     private HashMap<String, Integer> predicateIndex;
@@ -118,7 +119,7 @@ public class RandomRules {
                         }
                         if (subjectID != -99 && predicateID != -99 && objectID != -99) {
                             tripleHelp = new Triple(subjectID, predicateID,
-                                    objectID);
+                                    objectID, bodySubject, bodyPredicate, bodyObject);
                             body.add(tripleHelp);
                         }
                     }
@@ -152,7 +153,7 @@ public class RandomRules {
                     predicateID = -99;
                 }
                 if (subjectID != -99 && predicateID != -99 && objectID != -99) {
-                    head = new Triple(subjectID, predicateID, objectID);
+                    head = new Triple(subjectID, predicateID, objectID, headSubject, headPredicate, headObject);
                     rule = new Rule(head, body);
                 } else {
                     continue;
@@ -164,7 +165,7 @@ public class RandomRules {
 
                     if (rule.getBody().size() == 4) {
                         continuer = true;
-                        System.out.println(c++ + " Type2: " + rule);
+                        //System.out.println(c++ + " Type2: " + rule);
                     }
 
                     /*if (rule.getHead().getSubject() < 0
@@ -384,6 +385,7 @@ public class RandomRules {
         if (ruleSet != null) {
             filteredRules.addAll(ruleSet);
         }
+        //ilteredRules.forEach(e -> System.out.println(e.getId() + " : " + e));
         ArrayList<Integer> resultList;
         switch (Config.getStringValue("TESTRULES_METHOD")) {
             case "testRulesUnionAllShorterSelect":
@@ -439,7 +441,6 @@ public class RandomRules {
         int sub, pre, obj;
         try {
             BufferedReader reader = new BufferedReader(new FileReader(file));
-
             for (String line; (line = reader.readLine()) != null; ) {
                 importList = line.split("\\s+");
                 if (importList.length == 3) {
@@ -471,7 +472,7 @@ public class RandomRules {
[…]

### 2022-12-11 Code/sql-implementations/src/utils/Test.java

@@ -20,18 +20,18 @@ public class Test {
     Key2Int key2Int1 = new Key2Int(4, 16325);
     Key2Int key2Int2 = new Key2Int(4, 16325);
     Key2Int key2Int3 = new Key2Int(4, 16325);
-    Rule rule = new Rule(new Triple(-24, 4, 16325), new ArrayList<Triple>());
+    //Rule rule = new Rule(new Triple(-24, 4, 16325), new ArrayList<Triple>());
 
     if(!bothBound.containsKey(key2Int1)){
       bothBound.put(key2Int1, new ArrayList<>());
     }
-      bothBound.get(key2Int1).add(rule);
+      //bothBound.get(key2Int1).add(rule);
 
 
     if(!bothBound.containsKey(key2Int2)){
       bothBound.put(key2Int2, new ArrayList<>());
     }
-    bothBound.get(key2Int2).add(rule);
+    //bothBound.get(key2Int2).add(rule);
     System.out.println(bothBound.get(key2Int3));
   }
 }

### 2022-12-11 Paper/Overleaf/paper.tex

@@
 \begin{abstract}
-In this paper, we focus on the problem of finding all rules that imply a certain target fact given a knowledge graph and a set of previously learned rules over the knowledge graph.
-We specifically analyze how relational database technology performs in solving this problem using indexing, filtering and pre-processing methods.
-Our research might help to advance approaches for link prediction and can be used to efficiently explain the dynamics of rules given large amounts of target facts or large knowledge graphs.
-
-Test
+ Retrieving information from knowledge graphs is a highly relevant problem in various fields.
+This can for example be achieved by learning and applying rules based on the knowledge graphs.
+In this paper, we focus on the problem of identifying all rules that entail a certain target fact given a knowledge graph and a set of previously learned rules.
+We specifically aimed to create and analyze an efficient approach using relational database technology including indexing, filtering and pre-processing methods.
+This problem is relevant in the context of link prediction and explainability.
+Our experiments demonstrated the efficacy of our approach on different datasets like YAGO3-10, WN18RR and FB15k-237 using rules learned by the bottom up rule learner AnyBURL.
 \end{abstract}
 \begin{keywords}
-Knowledge graph completion \and Link prediction \and Knowledge representation and reasoning \and Relational databases.
+Knowledge graphs \and Relational databases \and Link prediction \and Explainability.
 \end{keywords}
@@
 Knowledge graphs (KGs) are used in different fields from biomedical applications \cite{OpenBioLink} over applications in the context of social networks \cite{SocialNetworks} to semantic search applications\cite{SemanticSearch}.
-We want to contribute by analyzing how to effectively check which rules of a KG imply certain target facts using relational database technology.
+Learned rules over the KG describe patterns of the KG and allow for knowledge inference.
+This can can for example be used to predict missing information also known as the link prediction problem.
+
+Identifying which previously learned rules of a KG entail certain target facts is an important problem for two main reasons.
+Firstly, it is relevant in the context of the link prediction problem.
+Prior research indicates that rule-based approaches and embedding models can be used in combination to improve inference quality.\cite{RuleEmbeddingCombination1}\cite{RuleEmbeddingCombination2} There is a significant need for a time-efficient approach which is used to check which rules entail certain facts in that context.
+Secondly, it is generally helpful to explain the connections between rules and (potential) facts.
+
+We want to contribute by creating and analyzing an effective approach to this problem using relational database technology.
+To achieve those goals, we ran several experiments aimed to illustrate the performance of our approach on different KGs (YAGO3-10, WN18RR and FB15k-237) and to show how different database setups impact the performance in terms of execution time using filtering, indexing and pre-processing methods.
 For the rule learning, we use AnyBURL, a fast bottom up rule learner for KGs.\cite{AnyBURL19}
 
-Our research might be used as a primitive for the link prediction problem.
-Prior research indicates that rule-based approaches and embedding models can be used in combination to improve inference quality.\cite{RuleEmbeddingCombination1}\cite{RuleEmbeddingCombination2} There is a significant need for a time-efficient solution that is used to check which rules implicate certain facts.
-Beyond potentially facilitating faster link prediction, our findings can already be helpful to explain the dynamics of rules given large amounts of potential facts for KGs.
-It might also particularly help in case of large knowledge graphs and memory constraints.
+\section{Preliminaries} 
+A KG is a set of (subject, relation, object)-triples also called facts.
+There is a set of entities present as subjects and objects and a set of relations in a KG.
+Here is an example KG with the entities \(peter\), \(anna\) and \(germany\) and the relations \(livesIn\) and \(marriedTo\).
 
-To achieve those goals, we ran several experiments aimed to illustrate how different database setups impact the performance in terms of execution time using filtering, indexing and pre-processing methods.
-Beyond that, we also performed a quantile performance analysis on the queries and rules.
+\[\textbf{KG} = \{marriedTo(anna, peter), bornIn(peter, germany), …\}\]
 
-\section{Preliminaries} 
-A knowledge graph is a set of triples (subject, relation, object) that describe a directed graph.
-The edges are represented by the relations going from subject to object.
-The nodes are the set of entities which appear as subjects and objects of the triples.
-The individual triples are also called facts.
-KGs can be used to describe various domains and to identify rules and patterns in those domains.
-For our purposes, we are concerned about first-order logic Horn rules.
-These rules consist of a head and a body, where the head consists of one triple and the body consists of one or more triples.
-The subjects and objects of the triples can either be specific entities or variables.
+For our purposes, we are concerned about first-order logic Horn rules, which describe patterns of KGs.
+These rules consist of a head and a body, where the head consists of one relation and the body consists of one or more relations.
 Here are two example rules.
-We use the notation relation(subject, object) for the triples and capitalize the variables.
+We capitalize the variables and lowercase the constants representing entities.
 
-\begin{equation}
+\begin{equation} \label{eq:rule_1}
 citizenOf(X, germany) \leftarrow bornIn(X, mannheim) 
-\end{equation}
+\end{equation} \label{eq:rule_2}
 \begin{equation}
@@
 
-The grounding of a rule is a set of triples of the KG that fulfil the rule, with matching entities for all variables.
-The triple set \textit{\{livesIn(tom, germany), marriedTo(tom, bob), bornIn(bob, germany)\}} would be a grounding for the second rule.
+A grounding of a rule assigns values to all variables of the rule, resulting in a ground term.
+The problem we are trying to solve is identifying rules that entail certain target facts which are not part of the KG based on the KG and a previously learned set of rules.
+This means we are interested in groundings where all body atoms are present in the KG and the head atom matches the target fact.
+To illustrate the problem, think of the following target fact.
 
-\section{Problem Statement}
-The problem we are trying to solve is identifying rules that imply certain target facts which are not part of the KG based on the KG and a previously learned set of rules.
-This means we are looking for a grounding of the rule with the target triple matching the head of the rule and a subset of triples from the KG matching the body of the rule.
-To illustrate the problem, think of the following example.
+\[\textbf{Target Fact} = livesIn(anna, germany)\]
 
-\begin{itemize}
-  \item \textbf{Knowledge Graph:}
-  \item[] \(\{…, marriedTo(anna, peter), bornIn(peter, germany), …\}\)
-  \item \textbf{Target Fact:} \(livesIn(anna, germany)\)
-  \item \textbf{Learned Rules:}
-  \begin{enumerate}
-      \item \(citizenOf(X, germany) \leftarrow bornIn(X, mannheim)\)
-      \item \(livesIn(X, germany) \leftarrow marriedTo(X, A_1), bornIn(A_1, germany)\)
-  \end{enumerate}
-\end{itemize}
-
-In our example we can start off with excluding rule one from further investigation as the target fact needs to match the head of the rule.
-Rule two has a matching head with \(X = anna\).
-To investigate whether rule two actually implies the target fact, we need to find matching facts for the rule body from the KG.
+In our example we can start off with excluding rule (\ref{eq:rule_1}) from further investigation as the target fact needs to match the head of the rule.
+Rule (\ref{eq:rule_2}) has a matching head with \(X = anna\).
+To investigate whether rule two actually entails the target fact, we need to find matching facts for the rule body from the KG.
 The KG contains the facts \(marriedTo(anna, peter)\) and \(bornIn(peter, germany)\).
@@
 
-\section{Implementation}
+\section{Proposed Approach}
 
-Our solution was implemented using the open-source object-relational database system PostgreSQL.
+Our approach was implemented using the open-source object-relational database system PostgreSQL.
 The basic idea behind our implementation is creating SQL queries for each target fact.
@@
 \section{Experiments}
-The goal of our experiments was to benchmark the performance of our solution and to measure the effects of different optimizations.
+The goal of our experiments was to benchmark the performance of our approach and to measure the effects of different optimizations.
 We also analyzed the execution time for individual queries and rules.
-Beyond that, we tested our solution on different datasets and rulesets.
-Our solution is written in Java using the PostgreSQL JDBC driver.
+Beyond that, we tested our approach on different datasets and rulesets.
+Our tool is written in Java using the PostgreSQL JDBC driver.
 The source code and datasets, used in the experiments, can be found at https://github.com/timgutberlet/From-Facts-to-Rules-using-Relational-Databases.
@@
 
+\subsection{Overall performance}
+
+\begin{table}[t]
+\centering
+\begin{tabular}{lllllll}
+\toprule
+ & XXXk rules & XXXk rules  & XXXk rules & \#entities & \#relations & \#facts\\
+\midrule
+YAGO3-10 & XX ms & 10 ms & XX ms & 123,182 & 37 & 1,179,040\\
+WN18RR & XX ms & XX ms & XX ms & 40,943 & 11 & 93,003\\
+FB15k-237 & XX ms & XX ms & XX ms & 14,505 & 237 & 310,079\\
+\bottomrule
+\end{tabular}
+\caption{Performance results (average execution time per target fact)}
+\label{tab:overall}
+\end{table}
+
+As illustrated in \Cref{tab:overall}, our approach works for different datasets and rulesets learned by AnyBURL in 10s, 50s and 100s.
+The approach is particularly well suited for datasets with many relations, like the Freebase dataset, as it allows for small B-trees of the indexes.
+
 \subsection{Ablation Study - Optimizations}
 
-\begin{table}
+\begin{table}[t]
 \centering
@@
 
+\begin{figure}[t]
 \begin{tikzpicture}
 \begin{axis}[
-        title={Fig. 1: Query quantile analysis (All optimizations enabled)},
 	x tick label style={
@@
 \end{tikzpicture}
+\caption{Query quantile analysis (All optimizations enabled)}
+\end{figure}
 
+\begin{figure}[t]
 \begin{tikzpicture}
 \begin{axis}[
-        title={Fig. 2: Rule quantile analysis (All optimizations enabled)},
 	x tick label style={
@@
 \end{tikzpicture}
+\caption{Rule quantile analysis (All optimizations enabled)}
+\end{figure}
 
@@
 The execution time for a specific rule is influenced by the length of the rule, as well as the target triple and the structure of the knowledge graph.
-
-\subsection{Further Datasets and Rulesets}
-
-\begin{table}
-\centering
-\begin{tabular}{llll}
-\toprule
- & AnyBURL - 10s & AnyBURL - 50s & AnyBURL - 100s\\
-\midrule
-YAGO3-10 & XX ms & 10 ms & XX ms\\
-WN18RR & XX ms & XX ms & XX ms\\
-FB15k-237 & XX ms & XX ms & XX ms\\
-\bottomrule
-\end{tabular}
-\caption{Average Execution Time for different Datasets and Rulesets}
-\label{tab:further-datasets}
-\end{table}
-
-Our solution works for different datasets and rulesets learned by AnyBURL in the given time.
-The solution is particularly well suited for datasets with many relations, like the Freebase dataset, as it allows for small B-trees of the indexes.
 
@@
 
-We demonstrated an effective solution for finding all rules that imply a certain target fact given a knowledge graph and a set of previously learned rules.
+We demonstrated an effective approach for finding all rules that imply a certain target fact given a knowledge graph and a set of previously learned rules.
 Our experiments specifically demonstrated the effect of filtering, indexing and pre-processing methods.
-Potential next steps include a further analysis of our solution on various datasets, a comparison of different database technologies (particularly triplestores), exploring the use of multithreading and creating a dedicated solution only using the main memory.
+Potential next steps include a further analysis of our approach on various datasets, a comparison of different database technologies (particularly triplestores), exploring the use of multithreading and creating a dedicated solution only using the main memory.
 

### 2022-12-11 Code/sql-implementations/src/database/DBFuncs.java
Fix error in dbFill

@@ -1528,47 +1528,66 @@ public class DBFuncs {
             long elapsedTime;
             long count = 0;
             Statement stmt = con.createStatement();
-            for (Integer integer : predicateList) {
-                sql = "DROP MATERIALIZED VIEW IF EXISTS v" + integer.toString() + " CASCADE;";
-                //stmt.addBatch(sql);
-                sql = "DROP MATERIALIZED VIEW IF EXISTS l" + integer.toString() + " CASCADE;";
-                //stmt.addBatch(sql);
-                sql = "DROP MATERIALIZED VIEW IF EXISTS r" + integer.toString() + " CASCADE;";
-                //stmt.addBatch(sql);
-                sql = "create materialized view  l" + integer.toString() + " as \n" +
-                        "select sub, obj from " + Config.getStringValue("KNOWLEDGEGRAPH_TABLE") + " where pre = " + integer + ";";
+            for(int i = 0; i < 200000; i++){
+                sql = "DROP MATERIALIZED VIEW IF EXISTS x" + i +";";
+                stmt.addBatch(sql);
                 count++;
-                //stmt.addBatch(sql);
-                sql = "CREATE UNIQUE INDEX l" + integer + "Index ON l" + integer + " (sub, obj);";
-                //stmt.addBatch(sql);
-                sql = "ALTER MATERIALIZED VIEW l" + integer + " CLUSTER ON l" + integer + "Index;";
-                //stmt.addBatch(sql);
-                //
-                sql = "DROP MATERIALIZED VIEW IF EXISTS r" + integer + ";";
-                //stmt.addBatch(sql);
-                sql = "create materialized view  r" + integer.toString() + " as \n" +
-                        "select sub, obj from " + Config.getStringValue("KNOWLEDGEGRAPH_TABLE") + " where pre = " + integer + ";";
-                count++;
-                //System.out.println(triple.toString());
-                //stmt.addBatch(sql);
-                sql = "CREATE UNIQUE INDEX r" + integer + "Index ON r" + integer + " (obj, sub);";
-                //stmt.addBatch(sql);
-                sql = "ALTER MATERIALIZED VIEW r" + integer + " CLUSTER ON r" + integer + "Index;";
-                //stmt.addBatch(sql);
-                //
-                /*if (count % 1 == 0 || count == predicateList.size()) {
+                if (count % 500 == 0){
                     stmt.executeBatch();
                     stmt.clearBatch();
                     elapsedTime = System.nanoTime() - startTime;
[…]

### 2022-12-11 Code/sql-implementations/src/database/DBFuncs.java
Fix error in dbFill 2

@@ -99,6 +99,100 @@ public class DBFuncs {
         //r.getBody().add(new Triple(-2, 3, -1));
         //createFunctionsBothUnbound(r);
     }
+    public static void setSubjectIndex(HashMap<String, Integer> map){
+        try {
+        PreparedStatement stmt;
+            String sql =
+                    "INSERT INTO subjects(id, txt) VALUES (?,?)";
+            stmt = con.prepareStatement(sql);
+            long startTime = System.nanoTime();
+            long elapsedTime;
+            long count = 0;
+        for(Map.Entry<String, Integer> entry : map.entrySet()){
+                    count++;
+                    stmt.setInt(1, entry.getValue());
+                    stmt.setString(2, entry.getKey());
+                    stmt.addBatch();
+                    if (count % 500 == 0 || count == map.size()) {
+                        stmt.executeBatch();
+                        stmt.clearBatch();
+                        elapsedTime = System.nanoTime() - startTime;
+                        startTime = System.nanoTime();
+                        System.out.println(
+                                "Inserted Subjects " + count + " of " + map.size() + " ; Time: " + (elapsedTime / 1000000)
+                                        + "ms");
+                    }
+        }
+        stmt.executeBatch();
+        con.commit();
+        stmt.close();
+        } catch (SQLException e) {
+            e.printStackTrace();
+        }
+    }
+    public static void setPredicateIndex(HashMap<String, Integer> map){
+        try {
+            PreparedStatement stmt;
+            String sql =
+                    "INSERT INTO predicates(id, txt) VALUES (?,?)";
+            stmt = con.prepareStatement(sql);
+            long startTime = System.nanoTime();
+            long elapsedTime;
+            long count = 0;
+            for(Map.Entry<String, Integer> entry : map.entrySet()){
+                count++;
+                stmt.setInt(1, entry.getValue());
+                stmt.setString(2, entry.getKey());
+                stmt.addBatch();
+                if (count % 500 == 0 || count == map.size()) {
+                    stmt.executeBatch();
+                    stmt.clearBatch();
+                    elapsedTime = System.nanoTime() - startTime;
+                    startTime = System.nanoTime();
+                    System.out.println(
+                            "Inserted Predicates " + count + " of " + map.size() + " ; Time: " + (elapsedTime / 1000000)
+                                    + "ms");
+                }
+            }
+            stmt.executeBatch();
+            con.commit();
+            stmt.close();
+        } catch (SQLException e) {
[…]

### 2022-12-11 Code/sql-implementations/src/utils/DbFill.java

@@ -192,7 +192,13 @@ public class DbFill {
         break;
       default:
     }
+    fillIndizes();
     System.out.println("Vocabulary filled");
+  }
+  public void fillIndizes() {
+    DBFuncs.setSubjectIndex(subjectIndex);
+    DBFuncs.setPredicateIndex(predicateIndex);
+    DBFuncs.setObjectIndex(objectIndex);
   }
 
   public void setIndexes() {

### 2022-12-11 Code/sql-implementations/src/database/DBFuncs.java
small dbFill fix

@@ -87,6 +87,23 @@ public class DBFuncs {
         }
     }
 
+    public static void deleteIndizes() {
+        Statement stmt;
+        String sql;
+        try {
+            sql = "DELETE FROM subjects";
+            stmt = con.createStatement();
+            stmt.executeQuery(sql);
+            sql = "DELETE FROM predicates";
+            stmt = con.createStatement();
+            stmt.executeQuery(sql);
+            sql = "DELETE FROM objects";
+            stmt = con.createStatement();
+            stmt.executeQuery(sql);
+        } catch (SQLException e) {
+        }
+    }
+
 
     /**
      * Used for testing

### 2022-12-11 Code/sql-implementations/src/utils/DbFill.java

@@ -196,6 +196,7 @@ public class DbFill {
     System.out.println("Vocabulary filled");
   }
   public void fillIndizes() {
+    DBFuncs.deleteIndizes();
     DBFuncs.setSubjectIndex(subjectIndex);
     DBFuncs.setPredicateIndex(predicateIndex);
     DBFuncs.setObjectIndex(objectIndex);

### 2022-12-11 Code/sql-implementations/src/database/DBFuncs.java
update JAR

@@ -83,23 +83,41 @@ public class DBFuncs {
             sql = "DELETE FROM " + Config.getStringValue("KNOWLEDGEGRAPH_TABLE");
             stmt = con.createStatement();
             stmt.executeQuery(sql);
+            con.commit();
         } catch (SQLException e) {
         }
     }
 
-    public static void deleteIndizes() {
+    public static void deleteSubjects() {
         Statement stmt;
         String sql;
         try {
-            sql = "DELETE FROM subjects";
+            sql = "DELETE FROM subjects;";
             stmt = con.createStatement();
             stmt.executeQuery(sql);
-            sql = "DELETE FROM predicates";
+            con.commit();
+        } catch (SQLException e) {
+        }
+    }
+    public static void deletePredicates() {
+        Statement stmt;
+        String sql;
+        try {
+            sql = "DELETE FROM predicates;";
             stmt = con.createStatement();
             stmt.executeQuery(sql);
-            sql = "DELETE FROM objects";
+            con.commit();
+        } catch (SQLException e) {
+        }
+    }
+    public static void deleteObjects() {
+        Statement stmt;
+        String sql;
+        try {
+            sql = "DELETE FROM objects;";
             stmt = con.createStatement();
             stmt.executeQuery(sql);
+            con.commit();
         } catch (SQLException e) {
         }
     }

### 2022-12-11 Code/sql-implementations/src/utils/DbFill.java

@@ -196,7 +196,9 @@ public class DbFill {
     System.out.println("Vocabulary filled");
   }
   public void fillIndizes() {
-    DBFuncs.deleteIndizes();
+    DBFuncs.deleteSubjects();
+    DBFuncs.deletePredicates();
+    DBFuncs.deleteObjects();
     DBFuncs.setSubjectIndex(subjectIndex);
     DBFuncs.setPredicateIndex(predicateIndex);
     DBFuncs.setObjectIndex(objectIndex);

### 2022-12-11 Code/sql-implementations/src/database/DBFuncs.java
Update Config

@@ -715,7 +715,7 @@ public class DBFuncs {
                     sql.append(sqlEnd);
                 }
             }
-            //System.out.println(sql.toString());
+            System.out.println(sql.toString());
             rs = stmt.executeQuery(sql.toString());
             while (rs.next()) {
                 if (rs.getString("?column?") != null) {

### 2022-12-11 Code/sql-implementations/src/database/DBFuncs.java
Update Config

@@ -715,7 +715,7 @@ public class DBFuncs {
                     sql.append(sqlEnd);
                 }
             }
-            System.out.println(sql.toString());
+            //System.out.println(sql.toString());
             rs = stmt.executeQuery(sql.toString());
             while (rs.next()) {
                 if (rs.getString("?column?") != null) {

### 2022-12-11 Code/sql-implementations/src/database/DBFuncs.java
Update Config

@@ -890,12 +890,12 @@ public class DBFuncs {
             sql.append(where);
             sql.append(sqlEnd);
 
-            //System.out.println(sql.toString());
+            System.out.println("Getting the time for a rule: " + sql.toString());
             String planningTime = "";
             String executionTime = "";
             rs = stmt.executeQuery(sql.toString());
             while (rs.next()) {
-                //System.out.println(rs.getString("QUERY PLAN"));
+                System.out.println(rs.getString("QUERY PLAN"));
                 if (rs.getString("QUERY PLAN").startsWith("Planning Time:")) {
                     planningTime = rs.getString("QUERY PLAN");
                 }
@@ -1116,9 +1116,11 @@ public class DBFuncs {
                 System.out.println(sql);
                 stmt.addBatch(sql.toString());
                 sql = new StringBuffer("CREATE UNIQUE INDEX x" + rule.getId() + "Index ON x" + rule.getId());
+                System.out.println("Subset, Objectset: " + subSet + ", " + objSet);
                 if (subSet && objSet) {
                     sql.append("(sub, obj);");
                     rule.setBound(0);
+                    System.out.println();
                 } else if (subSet && !objSet) {
                     sql.append("(sub);");
                     rule.setBound(1);
@@ -1128,6 +1130,7 @@ public class DBFuncs {
                 } else {
                     Debug.printMessage("Error in creating indizes for Optimized Materialized Rules");
                 }
+                System.out.println("Real Rule Bound : "+ rule.getBound());
                 stmt.addBatch(sql.toString());
                 sql = new StringBuffer("ALTER MATERIALIZED VIEW x" + rule.getId() + " CLUSTER ON x" + rule.getId() + "Index;");
                 stmt.addBatch(sql.toString());

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java

@@ -195,7 +195,7 @@ public class RandomRules {
                         }
                     }*/
 
-                    System.out.println(bodyCount++);
+                    //System.out.println(bodyCount++);
 
                     if (!continuer) {
                         continue;
@@ -251,7 +251,7 @@ public class RandomRules {
                 rule.setId(counter++);
                 System.out.println(counter);
             }
-            System.out.println("Import finished");
+            System.out.println("Rule Import finished");
             reader.close();
             if (Config.getStringValue("TESTRULES_METHOD").equals("testRulesSimpleViews") && Config.getStringValue("REFILL_TABLES").equals("YES")) {
                 DBFuncs.createNormalViewForRule(this);
@@ -697,11 +697,15 @@ public class RandomRules {
         if (ruleSet != null) {
             filteredRules.addAll(ruleSet);
         }
+        int i = 0;
         for (Rule rule : filteredRules) {
-            if (ruleTimeHashMap.containsKey(rule.getId())) {
-                ruleTimeHashMap.get(rule.getId()).addTime(DBFuncs.timePerRule(rule, triple));
-            } else {
-                ruleTimeHashMap.put(rule.getId(), new RuleTime(DBFuncs.timePerRule(rule, triple), rule));
+            if(rule.getBody().size() == 2){
+                System.out.println(i++);
+                if (ruleTimeHashMap.containsKey(rule.getId())) {
+                    ruleTimeHashMap.get(rule.getId()).addTime(DBFuncs.timePerRule(rule, triple));
+                } else {
+                    ruleTimeHashMap.put(rule.getId(), new RuleTime(DBFuncs.timePerRule(rule, triple), rule));
+                }
             }
         }
     }
@@ -711,15 +715,16 @@ public class RandomRules {
         ArrayList<RuleTime> helpList = new ArrayList<>();
         ArrayList<Rule> ruleList = new ArrayList<>();
         ArrayList<Double> timeList = new ArrayList<>();
-        ArrayList<Double> learnedTimelist = new ArrayList<>();
+        ArrayList<Double> learnedTimelist;
         int i = 0;
+        System.out.println("Rule Count: " + ruleCount);
         for (Map.Entry<Integer, RuleTime> entry : ruleTimeHashMap.entrySet()) {
             if(entry.getValue().getRule().getBody().size() == 2) {
                 i++;
                 helpList.add(entry.getValue());
-                timeList.add(entry.getValue().avg());
+                timeList.add(entry.getValue().sum());
[…]

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java
Update Config

@@ -14,6 +14,7 @@ import models.*;
  * @author timgutberlet
  */
 public class RandomRules {
+    private int ruleLearnCount = 0;
     private ArrayList<Triple> validationTriples = new ArrayList<>();
     private HashMap<Integer, RuleTime> ruleTimeHashMap = new HashMap<>();
     private HashMap<String, Integer> subjectIndex;
@@ -57,6 +58,7 @@ public class RandomRules {
 
     public void importRules() {
         Key3Int key3Int;
+        int atomCounter = 0;
         Key2Int key2Int;
         String file = Config.getStringValue("RULES_PATH");
         Rule rule;
@@ -249,7 +251,10 @@ public class RandomRules {
                 }
 
                 rule.setId(counter++);
-                System.out.println(counter);
+                //System.out.println(counter);
+                if(rule.getBody().size() == 2) {
+                    System.out.println("2 Atom rules " + atomCounter++);
+                };
             }
             System.out.println("Rule Import finished");
             reader.close();
@@ -697,10 +702,9 @@ public class RandomRules {
         if (ruleSet != null) {
             filteredRules.addAll(ruleSet);
         }
-        int i = 0;
         for (Rule rule : filteredRules) {
             if(rule.getBody().size() == 2){
-                System.out.println(i++);
+                System.out.println(ruleLearnCount++);
                 if (ruleTimeHashMap.containsKey(rule.getId())) {
                     ruleTimeHashMap.get(rule.getId()).addTime(DBFuncs.timePerRule(rule, triple));
                 } else {
@@ -724,7 +728,8 @@ public class RandomRules {
                 helpList.add(entry.getValue());
                 timeList.add(entry.getValue().sum());
             }
-            if(i == (ruleCount*0.05)){
+            //if(i == (ruleCount*0.05)){
+            if(i == 15){
                 break;
             }
         }
@@ -734,8 +739,8 @@ public class RandomRules {
             ruleList.add(ruleTime.getRule());
             preProcessedRules.put(ruleTime.getRule().getId(), ruleTime.getRule());
         }
-        rulePreSave(ruleList);
         DBFuncs.viewsForQuantiles(ruleList);
+        rulePreSave(ruleList);
         System.out.println("Quantil Calc old Rule learning");
         quantilCalcSum(timeList);
         printPriciestRuleTimes(timeList);

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java
Update minor mistake

@@ -58,7 +58,8 @@ public class RandomRules {
 
     public void importRules() {
         Key3Int key3Int;
-        int atomCounter = 0;
+        int twoAtomCounter = 0;
+        int threeAtomCounter = 0;
         Key2Int key2Int;
         String file = Config.getStringValue("RULES_PATH");
         Rule rule;
@@ -253,10 +254,14 @@ public class RandomRules {
                 rule.setId(counter++);
                 //System.out.println(counter);
                 if(rule.getBody().size() == 2) {
-                    System.out.println("2 Atom rules " + atomCounter++);
+                    System.out.println("2 Atom rules " + twoAtomCounter++);
+                };
+                if(rule.getBody().size() == 3) {
+                    System.out.println("3 Atom rules " + threeAtomCounter++);
                 };
             }
             System.out.println("Rule Import finished");
+            System.exit(0);
             reader.close();
             if (Config.getStringValue("TESTRULES_METHOD").equals("testRulesSimpleViews") && Config.getStringValue("REFILL_TABLES").equals("YES")) {
                 DBFuncs.createNormalViewForRule(this);

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java
Update JAR

@@ -261,7 +261,7 @@ public class RandomRules {
                 };
             }
             System.out.println("Rule Import finished");
-            System.exit(0);
+            //System.exit(0);
             reader.close();
             if (Config.getStringValue("TESTRULES_METHOD").equals("testRulesSimpleViews") && Config.getStringValue("REFILL_TABLES").equals("YES")) {
                 DBFuncs.createNormalViewForRule(this);

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java
Update JAR

@@ -612,7 +612,7 @@ public class RandomRules {
 
     public static void quantilCalcSum(ArrayList<Double> timeList) {
         Collections.sort(timeList);
-        timeList.forEach(aLong -> System.out.println(aLong));
+        //timeList.forEach(aLong -> System.out.println(aLong));
         double help;
         double quantile = 0.05;
         for (double i = quantile; i <= 1; i += quantile) {
@@ -753,7 +753,7 @@ public class RandomRules {
         System.out.println("Quantil Calc new Rule Learning");
         quantilCalcSum(learnedTimelist);
         printPriciestRuleTimes(learnedTimelist);
-        System.exit(0);
+        //System.exit(0);
     }
 
     public void printPriciestRuleTimes(ArrayList<Double> timeList){

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java
Update JAR

@@ -640,7 +640,7 @@ public class RandomRules {
         System.out.println("Average: " + ((double) average) + " ms");
         System.out.println("Min: " + ((double) min) + " ms");
         System.out.println("Max: " + ((double) max) + " ms");
-        System.exit(0);
+        //System.exit(0);
     }
 
     public static boolean isFloat(double value) {
@@ -753,7 +753,7 @@ public class RandomRules {
         System.out.println("Quantil Calc new Rule Learning");
         quantilCalcSum(learnedTimelist);
         printPriciestRuleTimes(learnedTimelist);
-        //System.exit(0);
+        System.exit(0);
     }
 
     public void printPriciestRuleTimes(ArrayList<Double> timeList){

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java
Update JAR

@@ -58,8 +58,11 @@ public class RandomRules {
 
     public void importRules() {
         Key3Int key3Int;
+        int oneAtomCounter = 0;
         int twoAtomCounter = 0;
         int threeAtomCounter = 0;
+        int fourAtomCounter = 0;
+        int fiveAtomCounter = 0;
         Key2Int key2Int;
         String file = Config.getStringValue("RULES_PATH");
         Rule rule;
@@ -253,14 +256,28 @@ public class RandomRules {
 
                 rule.setId(counter++);
                 //System.out.println(counter);
+                if(rule.getBody().size() == 1) {
+                    oneAtomCounter++;
+                };
                 if(rule.getBody().size() == 2) {
-                    System.out.println("2 Atom rules " + twoAtomCounter++);
+                    twoAtomCounter++;
                 };
                 if(rule.getBody().size() == 3) {
-                    System.out.println("3 Atom rules " + threeAtomCounter++);
+                    threeAtomCounter++;
+                };
+                if(rule.getBody().size() == 4) {
+                    fourAtomCounter++;
+                };
+                if(rule.getBody().size() == 5) {
+                    fiveAtomCounter++;
                 };
             }
-            System.out.println("Rule Import finished");
+                System.out.println("1 Atom rules " + oneAtomCounter);
+                System.out.println("2 Atom rules " + twoAtomCounter);
+                System.out.println("3 Atom rules " + threeAtomCounter);
+                System.out.println("4 Atom rules " + fourAtomCounter);
+                System.out.println("5 Atom rules " + fiveAtomCounter);
+                System.out.println("Rule Import finished");
             //System.exit(0);
             reader.close();
             if (Config.getStringValue("TESTRULES_METHOD").equals("testRulesSimpleViews") && Config.getStringValue("REFILL_TABLES").equals("YES")) {

### 2022-12-11 Code/sql-implementations/src/utils/RandomRules.java
Update JAR

@@ -335,44 +335,49 @@ public class RandomRules {
         ArrayList<Integer> resultList = new ArrayList<>();
         for (Map.Entry<Key2Int, ArrayList<Rule>> entry : objBound.entrySet()) {
             for (Rule r : entry.getValue()) {
-                filteredRules.add(r);
-                resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
-
+                if(r.getHead().getObject() == triple.getObject() && triple.getPredicate() == r.getHead().getPredicate()){
+                    filteredRules.add(r);
+                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
+                    filteredRules.clear();
+                }
             }
-            filteredRules.clear();
         }
         for (Map.Entry<Key2Int, ArrayList<Rule>> entry : subBound.entrySet()) {
             for (Rule r : entry.getValue()) {
-                filteredRules.add(r);
-                resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
-
+                if(r.getHead().getSubject() == triple.getSubject() && triple.getPredicate() == r.getHead().getPredicate()){
+                    filteredRules.add(r);
+                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
+                    filteredRules.clear();
+                }
             }
-            filteredRules.clear();
         }
         for (Map.Entry<Key3Int, ArrayList<Rule>> entry : bothBound.entrySet()) {
             for (Rule r : entry.getValue()) {
-                filteredRules.add(r);
-                resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
-
+                if(r.getHead().getObject() == triple.getObject() && r.getHead().getSubject() == triple.getSubject() && triple.getPredicate() == r.getHead().getPredicate()){
+                    filteredRules.add(r);
+                    resultList.addAll(DBFuncs.testRulesUnionAllShorterSelectViewsForRelations(filteredRules, triple));
+                    filteredRules.clear();
+                }
             }
-            filteredRules.clear();
         }
         for (Map.Entry<Integer, ArrayList<Rule>> entry : noBoundEqual.entrySet()) {
             for (Rule r : entry.getValue()) {
-                filteredRules.add(r);
[…]

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java
Update Files

@@ -544,7 +544,7 @@ public class RandomRules {
             //System.out.println(triple.toText());
             //resultMap.get(triple).getRuleList().forEach(e -> System.out.println(e.toString()));
             //System.out.println(elapsedTime / 1000000);
-            if (queries % 2 == 0) {
+            if (queries % 100 == 0) {
                 elapsedTime2 = System.nanoTime();
                 System.out.println("Gesamtzeit: " + ((elapsedTime2 - startTime2) / 1000000) + " ms");
                 System.out.println("Durchschnittszeit: " + (((elapsedTime2 - startTime2) / 1000000) / queries) + " ms");
@@ -755,8 +755,8 @@ public class RandomRules {
                 helpList.add(entry.getValue());
                 timeList.add(entry.getValue().sum());
             }
-            //if(i == (ruleCount*0.05)){
-            if(i == 15){
+            if(i == (ruleCount*0.05)){
+            //if(i == 15){
                 break;
             }
         }
@@ -769,10 +769,10 @@ public class RandomRules {
         DBFuncs.viewsForQuantiles(ruleList);
         rulePreSave(ruleList);
         System.out.println("Quantil Calc old Rule learning");
-        quantilCalcSum(timeList);
-        printPriciestRuleTimes(timeList);
         learnedTimelist = getLearnedRuleTime(preProcessedRules, this.validationTriples);
         System.out.println("Quantil Calc new Rule Learning");
+        quantilCalcSum(timeList);
+        printPriciestRuleTimes(timeList);
         quantilCalcSum(learnedTimelist);
         printPriciestRuleTimes(learnedTimelist);
         System.exit(0);

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java
Update config.properties

@@ -559,6 +559,7 @@ public class RandomRules {
         for (TimeTuple timeTuple : resultMap.values()) {
             timeList.add(timeTuple.getTime());
         }
+        System.out.println("Quantil Calc for overall Query Analysis");
         quantilCalc(timeList);
     }
 
@@ -602,7 +603,7 @@ public class RandomRules {
 
     public static void quantilCalc(ArrayList<Long> timeList) {
         Collections.sort(timeList);
-        timeList.forEach(aLong -> System.out.println(aLong));
+        //timeList.forEach(aLong -> System.out.println(aLong));
         double help;
         double quantile = 0.05;
         for (double i = quantile; i <= 1; i += quantile) {

### 2022-12-12 Paper/Overleaf/paper.tex
Paper - Implemented Feedback from Tim and reworked the Proposed Approach Section

@@
 This problem is relevant in the context of link prediction and explainability.
-Our experiments demonstrated the efficacy of our approach on different datasets like YAGO3-10, WN18RR and FB15k-237 using rules learned by the bottom up rule learner AnyBURL.
+Our experiments demonstrated the efficacy of our approach and the effect of various optimizations on different datasets like YAGO3-10, WN18RR and FB15k-237 using rules learned by the bottom up rule learner AnyBURL.
 \end{abstract}
@@
 We want to contribute by creating and analyzing an effective approach to this problem using relational database technology.
-To achieve those goals, we ran several experiments aimed to illustrate the performance of our approach on different KGs (YAGO3-10, WN18RR and FB15k-237) and to show how different database setups impact the performance in terms of execution time using filtering, indexing and pre-processing methods.
+To achieve those goals, we ran several experiments aimed to illustrate the performance of our approach on different KGs (YAGO3-10, WN18RR and FB15k-237).
+Moreover, we intend to show how different database setups impact the performance in terms of execution time using filtering, indexing and pre-processing methods.
 For the rule learning, we use AnyBURL, a fast bottom up rule learner for KGs.\cite{AnyBURL19}
@@
 For our purposes, we are concerned about first-order logic Horn rules, which describe patterns of KGs.
-These rules consist of a head and a body, where the head consists of one relation and the body consists of one or more relations.
 Here are two example rules.
@@
 
+These have a head (e.g., \(citizenOf(X, germany)\)) consisting of one relation and a body (e.g., \(bornIn(X, mannheim)\)) consisting of one or more relations.
 A grounding of a rule assigns values to all variables of the rule, resulting in a ground term.
@@
 
-\[\textbf{Target Fact} = livesIn(anna, germany)\]
+\[\textbf{Target fact} = livesIn(anna, germany)\]
 
 In our example we can start off with excluding rule (\ref{eq:rule_1}) from further investigation as the target fact needs to match the head of the rule.
-Rule (\ref{eq:rule_2}) has a matching head with \(X = anna\).
-To investigate whether rule two actually entails the target fact, we need to find matching facts for the rule body from the KG.
-The KG contains the facts \(marriedTo(anna, peter)\) and \(bornIn(peter, germany)\).
+Rule (\ref{eq:rule_2}) has a matching head with \(X = anna\) and entails the target fact if \(\exists A_1 (marriedTo(anna, A_1) \wedge bornIn(A_1, germany))\).The KG contains the facts \(marriedTo(anna, peter)\) and \(bornIn(peter, germany)\).
 Together with the target fact, this results in a grounding for the second rule with \(X = anna\) and \(A_1 = peter\).
-Therefore, rule two is part of the solution.
+Therefore, rule (\ref{eq:rule_2}) is part of the solution.
 
-\section{Proposed Approach}
+\section{Proposed approach}
 
 Our approach was implemented using the open-source object-relational database system PostgreSQL.
-The basic idea behind our implementation is creating SQL queries for each target fact.
-The given KG will be stored in one table of the relational database (columns: “sub” for subjects, “pre” for relations and “obj” for objects).
-The SQL statement consists of \textit{UNION ALL} operations over all sub-queries for the individual rules.
-The rules are saved in a list and in a loop over all rules, the sub-queries for rules with matching heads are created.
+The given KG is stored in one table of the relational database (\textit{kg\_table} with columns “sub” for subjects, “pre” for relations and “obj” for objects).
+The basic idea behind our implementation is creating SQL queries which check whether a certain rule entails a certain target fact.
+Those individual queries are then combined using \textit{UNION ALL} operations over all rules, where the head matches the target fact.
 To improve this basic idea, we employed several optimizations listed below.
-This would be the query for the example given in the problem statement:
+This would be the query for rule (\ref{eq:rule_2}) and the target fact given in the preliminaries:
 
-\textit{(\textbf{SELECT} rule\_1 \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} k0.sub = anna \textbf{AND} k0.obj = k1.sub \textbf{AND} k1.obj = germany \textbf{LIMIT} 1)  \textbf{UNION ALL} (...)}
+\textit{\textbf{SELECT} rule\_1 \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} k0.sub = anna \textbf{AND} k0.obj = k1.sub \textbf{AND} k1.obj = germany \textbf{LIMIT} 1}
 
-\subsection{Hash Maps for Rule Pre-filtering} 
+\subsection{Hash Maps for rule pre-filtering} 
 To speed up the pre-filtering of the rules, we sorted the given set of rules into a hash map with five keys based on the rule head type.
@@
 
-\subsection{Tables for each Relation}
+\subsection{Database structure}
 We created an alternative approach to having one big table for all facts of the KG.
-This approach uses one table for each relation in the KG therefore decreasing the individual table size.
+This approach uses \textit{one table for each relation} in the KG therefore decreasing the individual table size.
 They only contain two columns for the subject and the object.
@@
 
-\subsection{Indexing}
-To speed up the joining of the tables, we duplicated the knowledge graph tables and employed a unique clustered index, once with (subject, object) as key and once with (object, subject) as key.
-Within the SQL statement, the tables with the order (subject, object) are used in case of a fixed subject and the tables with the order (object, subject) are used in the case of a fixed object or no fixed subject and object.
-This ensures that an index is used as often as possible.
-The index speeds up the search and enables the direct access to the table contents through the B-trees.
+To speed up the search and enable direct access to the table contents through the B-trees, we duplicated the knowledge graph tables and employed a \textit{unique clustered index}, once with (subject, object) as key and once with (object, subject) as key.
+Within the generated SQL statements, the tables with the order (subject, object) are used in case of a fixed subject and the tables with the order (object, subject) are used in the case of a fixed object or no fixed subject and object.
+This ensures that the indexes are used as efficiently as possible.
 
@@
 
-\subsection{Ablation Study - Optimizations}
+\subsection{Ablation study - optimizations}
 
@@
 \toprule
-Experiment & Avg. Execution Time \\
+Experiment & Avg. execution time \\
 \midrule
@@
 \midrule
-Hash Maps for Rule Pre-filtering disabled & XX ms\\
+Rule pre-filtering disabled & XX ms\\
 Tables for each relation disabled & 140 ms\\
 Indexing disabled & 1680 ms\\
-Pre-processing of expensive rules disabled & 11 ms\\
+Pre-processing of expensive rules disabled & 18 ms\\
 

### 2022-12-12 Code/sql-implementations/src/database/DBFuncs.java
change ruleCounter

@@ -890,7 +890,7 @@ public class DBFuncs {
             sql.append(where);
             sql.append(sqlEnd);
 
-            System.out.println("Getting the time for a rule: " + sql.toString());
+            //System.out.println("Getting the time for a rule: " + sql.toString());
             String planningTime = "";
             String executionTime = "";
             rs = stmt.executeQuery(sql.toString());

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java

@@ -756,7 +756,7 @@ public class RandomRules {
                 helpList.add(entry.getValue());
                 timeList.add(entry.getValue().sum());
             }
-            if(i == (ruleCount*0.05)){
+            if(i == (ruleCount*0.01)){
             //if(i == 15){
                 break;
             }

### 2022-12-12 Code/sql-implementations/src/database/DBFuncs.java
change ruleCounter

@@ -895,7 +895,7 @@ public class DBFuncs {
             String executionTime = "";
             rs = stmt.executeQuery(sql.toString());
             while (rs.next()) {
-                System.out.println(rs.getString("QUERY PLAN"));
+                //System.out.println(rs.getString("QUERY PLAN"));
                 if (rs.getString("QUERY PLAN").startsWith("Planning Time:")) {
                     planningTime = rs.getString("QUERY PLAN");
                 }

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java

@@ -686,7 +686,7 @@ public class RandomRules {
         for (Triple triple : queryTriples) {
             learnQuerySearch(triple);
             queries++;
-            if (queries % 100 == 0) {
+            if (queries % 50 == 0) {
                 elapsedTime2 = System.nanoTime();
                 System.out.println("Gesamtzeit: " + ((elapsedTime2 - startTime2) / 1000000) + " ms");
                 System.out.println("Durchschnittszeit: " + (((elapsedTime2 - startTime2) / 1000000) / queries) + " ms");

### 2022-12-12 Code/sql-implementations/src/database/DBFuncs.java
JAR

@@ -1112,11 +1112,11 @@ public class DBFuncs {
                 preSelect.append(select);
                 preSelect.append(where);
                 sql.append(preSelect);
-                System.out.println(rule);
-                System.out.println(sql);
+                //System.out.println(rule);
+                //System.out.println(sql);
                 stmt.addBatch(sql.toString());
                 sql = new StringBuffer("CREATE UNIQUE INDEX x" + rule.getId() + "Index ON x" + rule.getId());
-                System.out.println("Subset, Objectset: " + subSet + ", " + objSet);
+                //System.out.println("Subset, Objectset: " + subSet + ", " + objSet);
                 if (subSet && objSet) {
                     sql.append("(sub, obj);");
                     rule.setBound(0);

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java

@@ -750,6 +750,7 @@ public class RandomRules {
         ArrayList<Double> learnedTimelist;
         int i = 0;
         System.out.println("Rule Count: " + ruleCount);
+        System.exit(0);
         for (Map.Entry<Integer, RuleTime> entry : ruleTimeHashMap.entrySet()) {
             if(entry.getValue().getRule().getBody().size() == 2) {
                 i++;

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java
JAR

@@ -749,15 +749,17 @@ public class RandomRules {
         ArrayList<Double> timeList = new ArrayList<>();
         ArrayList<Double> learnedTimelist;
         int i = 0;
+        double threshhold = (double) ruleCount * 0.01;
         System.out.println("Rule Count: " + ruleCount);
-        System.exit(0);
+        System.out.println("Threshhold " + threshhold);
+        //System.exit(0);
         for (Map.Entry<Integer, RuleTime> entry : ruleTimeHashMap.entrySet()) {
             if(entry.getValue().getRule().getBody().size() == 2) {
                 i++;
                 helpList.add(entry.getValue());
                 timeList.add(entry.getValue().sum());
             }
-            if(i == (ruleCount*0.01)){
+            if(i == (threshhold)){
             //if(i == 15){
                 break;
             }

### 2022-12-12 Code/sql-implementations/src/database/DBFuncs.java
small error fix

@@ -1135,7 +1135,7 @@ public class DBFuncs {
                 sql = new StringBuffer("ALTER MATERIALIZED VIEW x" + rule.getId() + " CLUSTER ON x" + rule.getId() + "Index;");
                 stmt.addBatch(sql.toString());
                 count++;
-                if (count % 2 == 0 || count == rules.size()) {
+                if (count % 5 == 0 || count == rules.size()) {
                     stmt.executeBatch();
                     stmt.clearBatch();
                     elapsedTime = System.nanoTime() - startTime;

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java

@@ -749,9 +749,9 @@ public class RandomRules {
         ArrayList<Double> timeList = new ArrayList<>();
         ArrayList<Double> learnedTimelist;
         int i = 0;
-        double threshhold = (double) ruleCount * 0.01;
-        System.out.println("Rule Count: " + ruleCount);
-        System.out.println("Threshhold " + threshhold);
+        int threshhold = (int) (ruleCount * 0.01);
+        //System.out.println("Rule Count: " + ruleCount);
+        //System.out.println("Threshhold " + threshhold);
         //System.exit(0);
         for (Map.Entry<Integer, RuleTime> entry : ruleTimeHashMap.entrySet()) {
             if(entry.getValue().getRule().getBody().size() == 2) {
@@ -759,11 +759,12 @@ public class RandomRules {
                 helpList.add(entry.getValue());
                 timeList.add(entry.getValue().sum());
             }
-            if(i == (threshhold)){
+            if(i >= (threshhold)){
             //if(i == 15){
                 break;
             }
         }
+        //System.out.println("Size of Helplist : " + helpList.size());
         Collections.sort(helpList);
         for(RuleTime ruleTime : helpList){
             ruleTime.getRule().setLearned();

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java
Change to get Quantil Analysis for all Rules

@@ -748,6 +748,7 @@ public class RandomRules {
         ArrayList<Rule> ruleList = new ArrayList<>();
         ArrayList<Double> timeList = new ArrayList<>();
         ArrayList<Double> learnedTimelist;
+        ArrayList<Double> timeListAll = new ArrayList<>();
         int i = 0;
         int threshhold = (int) (ruleCount * 0.01);
         //System.out.println("Rule Count: " + ruleCount);
@@ -759,6 +760,7 @@ public class RandomRules {
                 helpList.add(entry.getValue());
                 timeList.add(entry.getValue().sum());
             }
+            timeListAll.add(entry.getValue().sum());
             if(i >= (threshhold)){
             //if(i == 15){
                 break;
@@ -775,10 +777,10 @@ public class RandomRules {
         rulePreSave(ruleList);
         System.out.println("Quantil Calc old Rule learning");
         learnedTimelist = getLearnedRuleTime(preProcessedRules, this.validationTriples);
-        System.out.println("Quantil Calc new Rule Learning");
-        quantilCalcSum(timeList);
+        System.out.println("Quantil Calc new Rule Learning for all Rules");
+        quantilCalcSum(timeListAll);
         printPriciestRuleTimes(timeList);
-        quantilCalcSum(learnedTimelist);
+        //quantilCalcSum(learnedTimelist);
         printPriciestRuleTimes(learnedTimelist);
         System.exit(0);
     }

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java
Change to get Quantil Analysis for all Rules

@@ -782,7 +782,11 @@ public class RandomRules {
         printPriciestRuleTimes(timeList);
         //quantilCalcSum(learnedTimelist);
         printPriciestRuleTimes(learnedTimelist);
-        System.exit(0);
+        try {
+            Thread.sleep(30);
+        } catch (InterruptedException e) {
+            throw new RuntimeException(e);
+        }
     }
 
     public void printPriciestRuleTimes(ArrayList<Double> timeList){

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java
change in quantil Analysis

@@ -637,7 +637,7 @@ public class RandomRules {
         Collections.sort(timeList);
         //timeList.forEach(aLong -> System.out.println(aLong));
         double help;
-        double quantile = 0.05;
+        double quantile = 0.01;
         for (double i = quantile; i <= 1; i += quantile) {
             if (isFloat(i * timeList.size())) {
                 help = timeList.get((int) Math.floor(i * timeList.size()));
@@ -779,6 +779,7 @@ public class RandomRules {
         learnedTimelist = getLearnedRuleTime(preProcessedRules, this.validationTriples);
         System.out.println("Quantil Calc new Rule Learning for all Rules");
         quantilCalcSum(timeListAll);
+        System.exit(0);
         printPriciestRuleTimes(timeList);
         //quantilCalcSum(learnedTimelist);
         printPriciestRuleTimes(learnedTimelist);

### 2022-12-12 Code/sql-implementations/src/utils/RandomRules.java
update JAR

@@ -773,13 +773,13 @@ public class RandomRules {
             ruleList.add(ruleTime.getRule());
             preProcessedRules.put(ruleTime.getRule().getId(), ruleTime.getRule());
         }
+        quantilCalcSum(timeListAll);
+        System.exit(0);
         DBFuncs.viewsForQuantiles(ruleList);
         rulePreSave(ruleList);
         System.out.println("Quantil Calc old Rule learning");
         learnedTimelist = getLearnedRuleTime(preProcessedRules, this.validationTriples);
         System.out.println("Quantil Calc new Rule Learning for all Rules");
-        quantilCalcSum(timeListAll);
-        System.exit(0);
         printPriciestRuleTimes(timeList);
         //quantilCalcSum(learnedTimelist);
         printPriciestRuleTimes(learnedTimelist);

### 2022-12-12 Paper/Overleaf/paper.tex
Paper - DRAFT_v2

@@
 
-\pgfplotsset{width=12.5cm, height=4cm}
+\pgfplotsset{width=12.5cm, height=5.5cm}
 
@@
 \author[Tim Gutberlet \and Janik Sauerbier]
-{Tim Gutberlet\footnote{University of Mannheim, \email{tgutberl@mail.uni-mannheim.de}} \and
-Janik Sauerbier\footnote{University of Mannheim, \email{jsauerbi@mail.uni-mannheim.de}}}
+{Tim Gutberlet\footnote{University of Mannheim, \email{tim.gutberlet@students.uni-mannheim.de}} \and
+Janik Sauerbier\footnote{University of Mannheim, \email{janik.sauerbier@students.uni-mannheim.de}}}
 \startpage{1} % Beginn der Seitenzählung für diesen Beitrag / Start page
@@
 In this paper, we focus on the problem of identifying all rules that entail a certain target fact given a knowledge graph and a set of previously learned rules.
-We specifically aimed to create and analyze an efficient approach using relational database technology including indexing, filtering and pre-processing methods.
+We specifically aimed to create and analyze an efficient approach using relational database technology including indexing, filtering and pre-computing methods.
 This problem is relevant in the context of link prediction and explainability.
@@
 Learned rules over the KG describe patterns of the KG and allow for knowledge inference.
-This can can for example be used to predict missing information also known as the link prediction problem.
+This can can for example be used to predict missing or additional information also known as the link prediction problem.
 
@@
 To achieve those goals, we ran several experiments aimed to illustrate the performance of our approach on different KGs (YAGO3-10, WN18RR and FB15k-237).
-Moreover, we intend to show how different database setups impact the performance in terms of execution time using filtering, indexing and pre-processing methods.
+Moreover, we intend to show how different database setups impact the performance in terms of execution time using filtering, indexing and pre-computing methods.
 For the rule learning, we use AnyBURL, a fast bottom up rule learner for KGs.\cite{AnyBURL19}
@@
 The problem we are trying to solve is identifying rules that entail certain target facts which are not part of the KG based on the KG and a previously learned set of rules.
-This means we are interested in groundings where all body atoms are present in the KG and the head atom matches the target fact.
+This means we are interested in groundings where all body atoms exist in the KG and the head atom matches the target fact.
 To illustrate the problem, think of the following target fact.
@@
 
-\section{Proposed approach}
+\section{Proposed Approach}
 
 Our approach was implemented using the open-source object-relational database system PostgreSQL.
-The given KG is stored in one table of the relational database (\textit{kg\_table} with columns “sub” for subjects, “pre” for relations and “obj” for objects).
-The basic idea behind our implementation is creating SQL queries which check whether a certain rule entails a certain target fact.
+The given KG is stored in one table of the relational database (\textit{kg\_table} with columns “sub” for subjects, “rel” for relations and “obj” for objects).
+The basic idea behind our approach is creating SQL queries which check whether a certain rule entails a certain target fact.
 Those individual queries are then combined using \textit{UNION ALL} operations over all rules, where the head matches the target fact.
@@
 
-\textit{\textbf{SELECT} rule\_1 \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} k0.sub = anna \textbf{AND} k0.obj = k1.sub \textbf{AND} k1.obj = germany \textbf{LIMIT} 1}
+\textit{\textbf{SELECT} rule\_1 \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} k0.sub = anna \textbf{AND} k0.rel = marriedTo \textbf{AND} k0.obj = k1.sub \textbf{AND} k1.rel = bornIn \textbf{AND} k1.obj = germany \textbf{LIMIT} 1}
 
-\subsection{Hash Maps for rule pre-filtering} 
+\subsection{Advanced Rule Pre-filtering} 
 To speed up the pre-filtering of the rules, we sorted the given set of rules into a hash map with five keys based on the rule head type.
-Three of the head types have fixed entities, either as subject and/or as object [e.g., \textit{r1(e1, X)}, \textit{r1(X, e1)} or \textit{r1(e1, e2)}].
-Two of the head types have a variable as subject and as object, either the same variable or different variables [e.g., \textit{r1(X, X)}, \textit{r1(X, Y)}].
-Within the respective categories, the rules are again sorted into hash maps based on the relation and the fixed entities.
+Three of the head types have one or two fixed constants [e.g., \textit{r1(c1, X)}, \textit{r1(X, c1)} or \textit{r1(c1, c2)}].
+Two of the head types have two variables, either the same variable or different variables [e.g., \textit{r1(X, X)}, \textit{r1(X, Y)}].
+Within the respective categories, the rules are again sorted into hash maps based on the relation and the fixed constants.
 This allows for initial filtering of the rules in O(1).
 
-\subsection{Database structure}
-We created an alternative approach to having one big table for all facts of the KG.
-This approach uses \textit{one table for each relation} in the KG therefore decreasing the individual table size.
-They only contain two columns for the subject and the object.
-This solution improved the overall performance by reducing the size of the B-trees of the indexes.
+\subsection{Database Structure}
+Alternative to having one big table for all facts, our approach uses one table for each relation in the KG, with two columns for the subject and the object.
+This improved the performance by reducing the size of the B-trees of the indexes described in the following paragraph.
 
-To speed up the search and enable direct access to the table contents through the B-trees, we duplicated the knowledge graph tables and employed a \textit{unique clustered index}, once with (subject, object) as key and once with (object, subject) as key.
-Within the generated SQL statements, the tables with the order (subject, object) are used in case of a fixed subject and the tables with the order (object, subject) are used in the case of a fixed object or no fixed subject and object.
-This ensures that the indexes are used as efficiently as possible.
+To speed up the search and enable direct access to the table contents through the B-trees, we employed unique clustered indexes.
+We duplicated the knowledge graph tables and used once (subject, object) as key and once (object, subject) as key for the indexes.
+Within the generated SQL statements, the tables with the order (subject, object) are used in case of a fixed subject.
+The tables with the order (object, subject) are used in the case of a fixed object or no fixed subject and object.
+This ensures that the indexes are used efficiently.
 
-\subsection{Pre-processing of expensive Rules}
-As we will illustrate in our experiments, certain rules are way more expensive than others.
-To address that, we integrated a simple pre-processing method.
-We gather an independent set of n facts and used them as target facts.
-The queries were amended with the \textit{EXPLAIN ANALYZE} command to get the execution time for the individual rule sub-queries.
-Afterwards, we ranked all rules by the sum of the execution time over all queries.
-The top x\% of rules were then pre-processed, by calculating all potential rule heads which could form a grounding together with a set of triples from the KG as rule body.
-We excluded rules with one body triple, as pre-processing didn’t improve the performance.
-We also excluded rules with more than two body triples, as those rules incur a pre-processing time two orders of magnitude higher than rules with two body triples, while they only make up a small fraction of the most expensive and frequent rules.
+\subsection{Pre-computing of Expensive Rules}
+\label{pre-computation}
+As we will illustrate in our experiments, certain rules result in way more cost in terms of execution time than others.
+To address that, we pre-computed a specific portion of those rules.
+To identify the most expensive rules, we gathered an independent set of n target facts from the KG and ran their queries with the \textit{EXPLAIN ANALYZE} command to get the execution time for the individual rule sub-queries.
+Afterwards, we ranked all rules which appeared in the results by their average execution time multiplied with their number of appearances in the results.
+The top x\% of rules were then pre-computed, by calculating all potential assignments for variables in the rule head that form a grounding together with a set of facts from the KG.
+We excluded rules with one body atom, as pre-computing didn’t improve their performance.
+Moreover, we excluded rules with more than two body atoms, as those rules incur a pre-computing time two orders of magnitude higher than rules with two body atoms, while they only make up a small fraction of the most expensive and frequent rules.
 
 \section{Experiments}
-The goal of our experiments was to benchmark the performance of our approach and to measure the effects of different optimizations.
-We also analyzed the execution time for individual queries and rules.
-Beyond that, we tested our approach on different datasets and rulesets.
-Our tool is written in Java using the PostgreSQL JDBC driver.
-The source code and datasets, used in the experiments, can be found at https://github.com/timgutberlet/From-Facts-to-Rules-using-Relational-Databases.
+The goal of our experiments was to benchmark the performance of our approach on different datasets and rulesets, as well as to measure the effects of different optimizations.
+Beyond that, we analyzed the execution time for individual rules.
+The implementation of our approach was written in Java using the PostgreSQL JDBC driver.
+The source code and datasets, used for the experiments, can be found at https://github.com/timgutberlet/From-Facts-to-Rules-using-Relational-Databases.
 We conducted all experiments on a Fujitsu Esprimo P957 (construction year 2017) with 32 GB RAM, 512 GB SSD, an Intel i7-7700 @ 3.6 GHz CPU and Ubuntu 22.04.1 LTS as an operating system.
+The rule learning with AnyBURL was mostly done with the standard configuration of AnyBURL-22 available at https://web.informatik.uni-mannheim.de/AnyBURL/.
+We only extended the limit of body atoms from one to three for acyclic rules to match the cyclic rules, as we do not intent to discriminate between them.
 
-\subsection{Overall performance}
+\subsection{Overall Performance}
 
@@
 \toprule
- & XXXk rules & XXXk rules  & XXXk rules & \#entities & \#relations & \#facts\\
+ & 27k rules & 107k rules  & 167k rules & \#entities & \#relations & \#facts of KG\\
 \midrule
-YAGO3-10 & XX ms & 10 ms & XX ms & 123,182 & 37 & 1,179,040\\
-WN18RR & XX ms & XX ms & XX ms & 40,943 & 11 & 93,003\\
-FB15k-237 & XX ms & XX ms & XX ms & 14,505 & 237 & 310,079\\
+YAGO3-10 & 4 ms & 10 ms & 14 ms & 123,182 & 37 & 1,079,040\\
+WN18RR & 35 ms & 65 ms & 77 ms & 40,943 & 11 & 86,835\\
+FB15k-237 & 3 ms & 8 ms & 11 ms & 14,505 & 237 & 272,115\\
 \bottomrule
@@
 
-As illustrated in \Cref{tab:overall}, our approach works for different datasets and rulesets learned by AnyBURL in 10s, 50s and 100s.
-The approach is particularly well suited for datasets with many relations, like the Freebase dataset, as it allows for small B-trees of the indexes.
+As illustrated in \Cref{tab:overall}, our approach works for different datasets \cite{YAGO3}\cite{WN18RR}\cite{FB15k-237} and rulesets learned by AnyBURL in 10s (27k rules), 50s (107k rules) and 100s (167k rules).
+We used the training sets as the KGs and the test sets as the target triples (3k-20k triples).
+The validation sets (3k-20k triples) were used as target triples to create the rule rankings to pre-compute the top 1\% most expensive rules.
+Our approach is particularly well suited for datasets with many relations, like the Freebase dataset, as it allows more tables with smaller B-trees of the indexes.
 
-\subsection{Ablation study - optimizations}
+\subsection{Ablation Study - Optimizations}
 
@@
 \midrule
-Rule pre-filtering disabled & XX ms\\
-Tables for each relation disabled & 140 ms\\
-Indexing disabled & 1680 ms\\
-Pre-processing of expensive rules disabled & 18 ms\\
+Advanced rule pre-filtering disabled & 55 ms\\
+Tables for each relation disabled & 133 ms\\
+Indexing disabled & 1720 ms\\
+Pre-computing of expensive rules disabled & 16 ms\\
 
@@
 \end{tabular}
-\caption{Ablation Study of Optimizations using YAGO3-10}
+\caption{Ablation study of optimizations using YAGO3-10 \& 107k rules}
 \label{tab:ablation}
@@
 
-For these experiments, we used YAGO3-10 as the benchmark dataset.\cite{YAGO3} The dataset consists of a training set (1079040 triples), test set and validation set (5000 triples each).
-We used AnyBURL to learn rules for 50 seconds, only generating rules that do not have empty bodies (107301 Rules).
-We used the training set as the KG and the validation set as the target triples.
-For the \textit{Pre-processing of expensive Rules} optimization, we use the test set (n = 5000) to create the rule ranking and pre-process the top 1\% rules.
-The results in \Cref{tab:ablation} are based on executing all 5000 queries.
+For the ablation study in \Cref{tab:ablation}, we used YAGO3-10 as the benchmark dataset and the AnyBURL 50s ruleset (107k rules).
+The “tables for each relation” optimization specifically reduced the execution time for long rules.
+For rules with one body atom, the execution time was reduced by 12\%, for two body atoms by 25\% and for three body atoms by 74\%.
+This is caused by the increased use of the indexes during execution for longer rules.
 
-The \textit{Tables for each Relation} optimization specifically reduced the execution time for long rules.
-For rules with one body triple, the execution time was reduced by 12\%, for two body triples by 25\% and for three body triples by 74\%.
-This is caused by the increased use of the indexes during execution.
-
-\subsection{Quantile Performance Analysis}
+\subsection{Rule Quantile Performance Analysis}
 
@@
 		/pgf/number format/1000 sep=},
-	ylabel=Execution Time (ET) in ms,
+	ylabel=\#appearances * execution time (ms),
         xlabel=Percentile,
@@
 	anchor=north,legend columns=-1},
-	ybar interval=0.5,
+	ybar interval=0.3,
+        axis x discontinuity=crunch
 ]
+
 \addplot 
-	coordinates {(0, 0.984156) (5, 4.113395)
-            (10, 4.242852) (15, 4.38007)
-            (20, 4.5406935) (25, 4.6855745)
-            (30, 4.860181) (35, 5.1234675)
-            (40, 5.626696) (45, 8.143013)
-            (50, 8.648211) (55, 8.905306)
-            (60, 9.140784) (65, 9.355278)
-            (70, 9.636537) (75, 9.959422)
-            (80, 10.475147) (85, 17.016464)
-            (90, 30.580482) (95, 32.047903) 
-            (100, 0)};
+	coordinates {(80, 3.909)
+            (81, 4.361) (82, 4.934)
+            (83, 6.515) (84, 10.682)
+            (85, 16.204) (86, 17.215)
+            (87, 23.746) (88, 28.900)
+            (89, 32.942) (90, 38.231)
+            (91, 49.141) (92, 60.599)
+            (93, 71.221) (94, 135.862)
+            (95, 235.4) (96, 348.644)
+            (97, 352.313) (98, 367.095)
+            (99, 374.032) (100, 1731.569)
+            (101, 0)};
 \end{axis}
-\label{Test}
 \end{tikzpicture}
-\caption{Query quantile analysis (All optimizations enabled)}
+\caption{Quantile performance analysis of rules using YAGO3-10 \& 107k rules}
+\label{fig:quantile-analysis}
 \end{figure}
 
-\begin{figure}[t]
-\begin{tikzpicture}
-\begin{axis}[
-	x tick label style={
-		/pgf/number format/1000 sep=},
-	ylabel=1/Frequency * Avg. ET in ms,
-        xlabel=Percentile,
-	enlargelimits=0.05,
-	legend style={at={(0.5,-0.1)},
-	anchor=north,legend columns=-1},
-	ybar interval=0.5,
-]
-\addplot 
-	coordinates {(0, 0.021) (5, 0.025)
-            (10, 0.026) (15, 0.027)
-            (20, 0.027) (25, 0.028)
-            (30, 0.030) (35, 0.031)
-            (40, 0.032) (45, 0.036)
-            (50, 0.042) (55, 0.052)
-            (60, 0.057) (65, 0.064)
-            (70, 0.08) (75, 0.1)
-            (80, 0.152) (85, 1.64)
-            (90, 8.26) (95, 12.90) 
-            (100, 0) };
-\end{axis}
-\label{Test}
-\end{tikzpicture}
-\caption{Rule quantile analysis (All optimizations enabled)}
-\end{figure}
-
-The analysis indicates that only few rules and few queries incur a major share of the cost.
-The \textit{Pre-processing of expensive Rules} optimization helped to reduce the cost for very expensive rules.
-The execution time for a specific rule is influenced by the length of the rule, as well as the target triple and the structure of the knowledge graph.
-
-%However, to show the relevance of our approach, we also tested the performance of our optimisation strategies, without limiting the rule types in any sense.
-
-%Conclusions / Assumptions:
-%Yago3 - 10 has relatively low amount of 37 different relations.
-Working with datasets with a higher count of relations could possibly improve the approach of using materialized views for every relation, as the knowledgegraph would be split even more fine-grained.
+The quantile analysis in \Cref{fig:quantile-analysis} is based on the number of appearances multiplied with the average execution of the rules after using the validation set as target triples as described in \ref{pre-computation}.
+It indicates that only few rules incur a major share of the cost.
+This ranking is used to determine the top x\% for the “pre-computing of expensive rules” optimization.
+In the given example (YAGO3-10 \& 107k rules), we achieved an 98.8\% execution time reduction for the pre-computed top 1\% rules.
+The pre-compution took five minutes.
+This reduced our average execution time from 16 ms to 10 ms, as illustrated in \Cref{tab:ablation}.
 
@@
 We demonstrated an effective approach for finding all rules that imply a certain target fact given a knowledge graph and a set of previously learned rules.
-Our experiments specifically demonstrated the effect of filtering, indexing and pre-processing methods.
+Our experiments specifically demonstrated the effect of filtering, indexing and pre-computing methods.
 Potential next steps include a further analysis of our approach on various datasets, a comparison of different database technologies (particularly triplestores), exploring the use of multithreading and creating a dedicated solution only using the main memory.

### 2022-12-16 Paper/Overleaf/paper.tex
Paper - Final Version

@@
 
-\pgfplotsset{width=12.5cm, height=5.5cm}
+\pgfplotsset{width=12.5cm, height=5.3cm}
 
@@
 %%% Das optionale Argument (sofern angegeben) wird für die Kopfzeile verwendet.
-\title[From Facts to Rules using Relational Databases]{From Facts to Rules using Relational Databases}
+\title[Which Rules Entail this Fact? An Efficient Approach Using RDBMSs]{Which Rules Entail this Fact? An Efficient Approach Using RDBMSs}
 %%% \subtitle{From Facts to Rules using Relational Databases} % if needed
@@
 \startpage{1} % Beginn der Seitenzählung für diesen Beitrag / Start page
-\editor{K.-U. Sattler et al.} % Names of Editors
+\editor{B. König-Ries et al.} % Names of Editors
 \booktitle{Datenbanksysteme für Business, Technologie und Web (BTW 2023)} % Name of book title
@@
 \begin{abstract}
- Retrieving information from knowledge graphs is a highly relevant problem in various fields.
-This can for example be achieved by learning and applying rules based on the knowledge graphs.
  In this paper, we focus on the problem of identifying all rules that entail a certain target fact given a knowledge graph and a set of previously learned rules.
-We specifically aimed to create and analyze an efficient approach using relational database technology including indexing, filtering and pre-computing methods.
 This problem is relevant in the context of link prediction and explainability.
-Our experiments demonstrated the efficacy of our approach and the effect of various optimizations on different datasets like YAGO3-10, WN18RR and FB15k-237 using rules learned by the bottom up rule learner AnyBURL.
+We propose an efficient approach using relational database technology including indexing, filtering and pre-computing methods.
+Our experiments demonstrate the efficiency of our approach and the effect of various optimizations on different datasets like YAGO3-10, WN18RR and FB15k-237 using rules learned by the bottom up rule learner AnyBURL.
+
 \end{abstract}
@@
 \section{Introduction}
-Knowledge graphs (KGs) are used in different fields from biomedical applications \cite{OpenBioLink} over applications in the context of social networks \cite{SocialNetworks} to semantic search applications\cite{SemanticSearch}.
+Knowledge graphs (KGs) are used in different fields from biomedical applications \cite{OpenBioLink} over applications in the context of social networks \cite{SocialNetworks} to semantic search applications \cite{SemanticSearch}.
 Learned rules over the KG describe patterns of the KG and allow for knowledge inference.
-This can can for example be used to predict missing or additional information also known as the link prediction problem.
+This can be for example used to predict missing or additional information also known as the link prediction problem.
 
 Identifying which previously learned rules of a KG entail certain target facts is an important problem for two main reasons.
-Firstly, it is relevant in the context of the link prediction problem.
-Prior research indicates that rule-based approaches and embedding models can be used in combination to improve inference quality.\cite{RuleEmbeddingCombination1}\cite{RuleEmbeddingCombination2} There is a significant need for a time-efficient approach which is used to check which rules entail certain facts in that context.
-Secondly, it is generally helpful to explain the connections between rules and (potential) facts.
+Firstly, to determine the confidence of an (unknown) fact, one must know from which rules it can be derived.
+The confidence of a fact refers to the probability that a particular fact within the graph is correct.
+Knowing the confidence is needed for rule-based link prediction models \cite{AnyBURL19}.
+Moreover, the confidence is used for building ensembles with embedding models \cite{RuleEmbeddingCombination1} and for improving embedding models \cite{RuleEmbeddingCombination2}.
+Secondly, it is generally helpful to be able to explain the connections between rules and (potential) facts.
+There is a significant need for a time-efficient approach to solve this problem.
 
-We want to contribute by creating and analyzing an effective approach to this problem using relational database technology.
-To achieve those goals, we ran several experiments aimed to illustrate the performance of our approach on different KGs (YAGO3-10, WN18RR and FB15k-237).
-Moreover, we intend to show how different database setups impact the performance in terms of execution time using filtering, indexing and pre-computing methods.
-For the rule learning, we use AnyBURL, a fast bottom up rule learner for KGs.\cite{AnyBURL19}
+In this paper, we create and analyze an effective approach to this problem, using relational database technology.
+Our approach uses indexing, filtering and pre-computing methods tailored to the natural structure of a KG and its respective rules.
+To illustrate the performance of our approach, we ran several experiments on different KGs (YAGO3-10 \cite{YAGO3}, WN18RR \cite{WN18RR} and FB15k-237 \cite{FB15k-237}).
+Moreover, we show how different database setups impact the performance for different rule lengths.
+For the rule learning, we use AnyBURL, a fast bottom up rule learner for KGs \cite{AnyBURL19}.
 
@@
 A KG is a set of (subject, relation, object)-triples also called facts.
-There is a set of entities present as subjects and objects and a set of relations in a KG.
+There is a set of entities present as subjects and objects as well as a set of relations in a KG.
 Here is an example KG with the entities \(peter\), \(anna\) and \(germany\) and the relations \(livesIn\) and \(marriedTo\).
 
-\[\textbf{KG} = \{marriedTo(anna, peter), bornIn(peter, germany), …\}\]
+\begin{equation*} \label{eq:knowledgegraph}
+\textbf{KG} = \{(anna,marriedTo, peter), (peter,bornIn,germany)\}\end{equation*}
 
@@
 \begin{equation} \label{eq:rule_1}
-citizenOf(X, germany) \leftarrow bornIn(X, mannheim) 
-\end{equation} \label{eq:rule_2}
-\begin{equation}
-livesIn(X, germany) \leftarrow marriedTo(X, A_1), bornIn(A_1, germany)
+(X, citizenOf, germany) \leftarrow (X, bornIn, mannheim) 
+\end{equation} 
+\begin{equation} \label{eq:rule_2}
+(X, livesIn, germany) \leftarrow (X, marriedTo, A_1), (A_1, bornIn, germany)
 \end{equation}
 
-These have a head (e.g., \(citizenOf(X, germany)\)) consisting of one relation and a body (e.g., \(bornIn(X, mannheim)\)) consisting of one or more relations.
-A grounding of a rule assigns values to all variables of the rule, resulting in a ground term.
-The problem we are trying to solve is identifying rules that entail certain target facts which are not part of the KG based on the KG and a previously learned set of rules.
-This means we are interested in groundings where all body atoms exist in the KG and the head atom matches the target fact.
+Each rule has a head (e.g., \((X, citizenOf, germany)\)) consisting of one atom and a body (e.g., \((X, 
+bornIn, mannheim)\)) consisting of one or more atoms.
+A grounding of a rule assigns values to all variables of the rule, resulting in a ground rule.
+A true body grounding refers to a grounding of the body of a rule for which all (ground) atoms appear in the KG.
+The problem we are solving is identifying all rules that entail certain target facts based on the KG and a previously learned set of rules.
+This means we are interested in whether there exists a true body grounding and the head atom can be unified with the target fact.
 To illustrate the problem, think of the following target fact.
 
-\[\textbf{Target fact} = livesIn(anna, germany)\]
+\begin{equation*} \label{eq:target_1}
+\textbf{Target fact} = (anna, livesIn, germany)
+\end{equation*}
 
-In our example we can start off with excluding rule (\ref{eq:rule_1}) from further investigation as the target fact needs to match the head of the rule.
-Rule (\ref{eq:rule_2}) has a matching head with \(X = anna\) and entails the target fact if \(\exists A_1 (marriedTo(anna, A_1) \wedge bornIn(A_1, germany))\).The KG contains the facts \(marriedTo(anna, peter)\) and \(bornIn(peter, germany)\).
-Together with the target fact, this results in a grounding for the second rule with \(X = anna\) and \(A_1 = peter\).
+Rule (\ref{eq:rule_1}) does not entail this fact because its head atom cannot be unified with the target fact.
+The head of rule (\ref{eq:rule_2}) can be unified with the target fact by assigning \(X = anna\).
+Therefore, it would entail the target fact if \(\exists A_1 ((anna, marriedTo, A_1) \wedge (A_1, bornIn, germany))\).
+The KG contains the facts \((anna, marriedTo, peter)\) and \((peter, bornIn, germany)\).
+Together with the target fact, this results in a true body grounding for rule (\ref{eq:rule_2}) with \(X = anna\) and \(A_1 = peter\).
 Therefore, rule (\ref{eq:rule_2}) is part of the solution.
@@
 
-Our approach was implemented using the open-source object-relational database system PostgreSQL.
 The given KG is stored in one table of the relational database (\textit{kg\_table} with columns “sub” for subjects, “rel” for relations and “obj” for objects).
 The basic idea behind our approach is creating SQL queries which check whether a certain rule entails a certain target fact.
-Those individual queries are then combined using \textit{UNION ALL} operations over all rules, where the head matches the target fact.
-To improve this basic idea, we employed several optimizations listed below.
+Those individual queries are then combined using \textit{UNION ALL} operations over all rules, where the head can unify with the target fact.
+To improve this basic idea, we employ several optimizations listed below.
 This would be the query for rule (\ref{eq:rule_2}) and the target fact given in the preliminaries:
 
-\textit{\textbf{SELECT} rule\_1 \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} k0.sub = anna \textbf{AND} k0.rel = marriedTo \textbf{AND} k0.obj = k1.sub \textbf{AND} k1.rel = bornIn \textbf{AND} k1.obj = germany \textbf{LIMIT} 1}
-
-\subsection{Advanced Rule Pre-filtering} 
-To speed up the pre-filtering of the rules, we sorted the given set of rules into a hash map with five keys based on the rule head type.
-Three of the head types have one or two fixed constants [e.g., \textit{r1(c1, X)}, \textit{r1(X, c1)} or \textit{r1(c1, c2)}].
-Two of the head types have two variables, either the same variable or different variables [e.g., \textit{r1(X, X)}, \textit{r1(X, Y)}].
-Within the respective categories, the rules are again sorted into hash maps based on the relation and the fixed constants.
-This allows for initial filtering of the rules in O(1).
+\textit{\textbf{SELECT} rule\_1 \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} t0.sub = anna \textbf{AND} t0.rel = marriedTo \textbf{AND} t0.obj = t1.sub \textbf{AND} t1.rel = bornIn \textbf{AND} t1.obj = germany \textbf{LIMIT} 1;}
 
 \subsection{Database Structure}
-Alternative to having one big table for all facts, our approach uses one table for each relation in the KG, with two columns for the subject and the object.
-This improved the performance by reducing the size of the B-trees of the indexes described in the following paragraph.
+\label{database-structure}
+Alternative to having one big table for all facts, our approach uses one table for each relation in the KG, with two columns for the subjects and the objects.
+This improves the performance by reducing the size of the B-trees of the indexes described in the following paragraph.
 
-To speed up the search and enable direct access to the table contents through the B-trees, we employed unique clustered indexes.
-We duplicated the knowledge graph tables and used once (subject, object) as key and once (object, subject) as key for the indexes.
+To speed up the search and enable direct access to the table contents through B-trees, we employ unique clustered indexes.
+We duplicate the knowledge graph tables and use once (subject, object) as key and once (object, subject) as key for the indexes.
 Within the generated SQL statements, the tables with the order (subject, object) are used in case of a fixed subject.
@@
 
+\subsection{Advanced Rule Pre-filtering} 
+Firstly, we only consider rules that can unify with the target fact for the query.
+This can be done by naively testing each rule.
+However, it can also be done more efficiently.
+The easiest way is storing each rule under its head as key in a hash map.
+The head stays unchanged, but we use “X” and “Y” as variable descriptors.
+
+Let (s, r, o) be a target triple with the entities s and o and the relation r.
+Then we only use the rules stored under the keys \textit{(X, r, Y)}, \textit{(s, r, X)}, \textit{(X, r, o)} or \textit{(s, r, o)} for the query.
+Additionally, when it is true that s=o for the target, we also use the key \textit{(X, r, X)}.
+Instead of looping through n rules naively in O(n), this allows for initial filtering of the rules in O(1).
+
 \subsection{Pre-computing of Expensive Rules}
@@
 As we will illustrate in our experiments, certain rules result in way more cost in terms of execution time than others.
-To address that, we pre-computed a specific portion of those rules.
-To identify the most expensive rules, we gathered an independent set of n target facts from the KG and ran their queries with the \textit{EXPLAIN ANALYZE} command to get the execution time for the individual rule sub-queries.
-Afterwards, we ranked all rules which appeared in the results by their average execution time multiplied with their number of appearances in the results.
-The top x\% of rules were then pre-computed, by calculating all potential assignments for variables in the rule head that form a grounding together with a set of facts from the KG.
-We excluded rules with one body atom, as pre-computing didn’t improve their performance.
-Moreover, we excluded rules with more than two body atoms, as those rules incur a pre-computing time two orders of magnitude higher than rules with two body atoms, while they only make up a small fraction of the most expensive and frequent rules.
+To address that, we pre-compute a portion of those rules.
+To identify the most expensive rules, we gather an independent set of n target facts from the KG and run their queries with the \textit{EXPLAIN ANALYZE} command to get the execution time for the individual rule sub-queries.
+Afterwards, we rank all rules which appear in the results by their average execution time multiplied with their number of appearances in the results.
+
+The top x\% of rules are then pre-computed, by calculating all potential assignments for variables in the rule head that form a grounding together with a set of facts from the KG.
+We then store these combinations in a table for the rule.
+If the subject or the object is a variable and the other one a constant, then the table only contains the column for the variable.
+For the implementation of the tables, we use materialized views to \(SELECT\) from the KG with all the conditions we already know from the rule.
+Consequently, the \(CREATE\) statement for pre-computing rule (\ref{eq:rule_2}) and the query for the target fact from the preliminaries would be:
+
+\textit{\textbf{CREATE MATERIALIZED VIEW} view\_rule\_1 \textbf{SELECT} t0.sub \textbf{AS} sub \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} t0.rel = marriedTo \textbf{AND} t0.obj = t1.sub \textbf{AND} t1.rel = bornIn \textbf{AND} t1.obj = germany};
+
+\textit{\textbf{SELECT} rule\_1 \textbf{FROM} view\_rule\_1  \textbf{WHERE} sub = anna \textbf{LIMIT} 1};
+
+We exclude rules with one body atom, as pre-computing does not improve their performance.
+Moreover, we exclude rules with more than two body atoms, as those rules incur a pre-computing time two orders of magnitude higher than rules with two body atoms, while they only make up a fraction of the most expensive and frequent rules.
 
 \section{Experiments}
-The goal of our experiments was to benchmark the performance of our approach on different datasets and rulesets, as well as to measure the effects of different optimizations.
-Beyond that, we analyzed the execution time for individual rules.
-The implementation of our approach was written in Java using the PostgreSQL JDBC driver.
-The source code and datasets, used for the experiments, can be found at https://github.com/timgutberlet/From-Facts-to-Rules-using-Relational-Databases.
+The goal of our experiments is to benchmark the performance of our approach on different datasets and rulesets, as well as to measure the effects of different optimizations.
+Beyond that, we analyze the execution time for individual rules.
+
+For the implementation of our approach, we used the open-source object-relational database system PostgreSQL.
+Additionally, we used Java with the PostgreSQL JDBC driver.
+The source code and datasets, we used for the experiments, can be found at https://github.com/timgutberlet/Which-Rules-Entail-This-Fact.
 We conducted all experiments on a Fujitsu Esprimo P957 (construction year 2017) with 32 GB RAM, 512 GB SSD, an Intel i7-7700 @ 3.6 GHz CPU and Ubuntu 22.04.1 LTS as an operating system.
 The rule learning with AnyBURL was mostly done with the standard configuration of AnyBURL-22 available at https://web.informatik.uni-mannheim.de/AnyBURL/.
-We only extended the limit of body atoms from one to three for acyclic rules to match the cyclic rules, as we do not intent to discriminate between them.
+We only extended the limit of body atoms from one to three for acyclic rules to match the cyclic rules, as we do not intend to discriminate between them.
 
@@
 \centering
-\begin{tabular}{lllllll}
+\begin{tabular}{l|lll|lll}
 \toprule
- & 27k rules & 107k rules  & 167k rules & \#entities & \#relations & \#facts of KG\\
+ & 10s & 50s  & 100s & \#entities & \#relations & \#facts of KG\\
 \midrule
@@
 WN18RR & 35 ms & 65 ms & 77 ms & 40,943 & 11 & 86,835\\
-FB15k-237 & 3 ms & 8 ms & 11 ms & 14,505 & 237 & 272,115\\
+FB15k-237 & 5 ms & 26 ms & 43 ms & 14,505 & 237 & 272,115\\
 \bottomrule
 \end{tabular}
-\caption{Performance results (average execution time per target fact)}
+\caption{Performance results (avg. execution time per target fact | dataset information)}
 \label{tab:overall}
@@
 
-As illustrated in \Cref{tab:overall}, our approach works for different datasets \cite{YAGO3}\cite{WN18RR}\cite{FB15k-237} and rulesets learned by AnyBURL in 10s (27k rules), 50s (107k rules) and 100s (167k rules).
-We used the training sets as the KGs and the test sets as the target triples (3k-20k triples).
-The validation sets (3k-20k triples) were used as target triples to create the rule rankings to pre-compute the top 1\% most expensive rules.
-Our approach is particularly well suited for datasets with many relations, like the Freebase dataset, as it allows more tables with smaller B-trees of the indexes.
+\begin{table}[t]
+\centering
+\begin{tabular}{l|lll|lll}
+\toprule
+ & 10s & 50s & 100s & 1 body atom & 2 body atoms & 3 body atoms\\
+\midrule
+YAGO3-10 & 27k & 107k & 165k & 82.8\% & 11.2\% & 6.0\%\\
+WN18RR & 6k & 20k & 30k & 33.0\% & 23.9\% & 43.1\%\\
+FB15k-237 & 33k & 135k & 252k & 53.9\% & 26.2\% & 19.9\%\\
+\bottomrule
+\end{tabular}
+\caption{Composition of the rules (\# of rules | avg. share by \# of body atoms)}
+\label{tab:rules}
+\end{table}
+
+As illustrated in \Cref{tab:overall}, our approach works for different datasets and rulesets learned by AnyBURL.
+For every dataset, we learned rules for 10s, 50s, and 100s (\Cref{tab:rules}).
+We use the training sets as the KGs and the test sets as the target triples (3k-20k triples).
+The validation sets (3k-20k triples) are used as target triples to create the rule rankings to pre-compute the top 1\% most expensive rules.
 
@@
 
-For the ablation study in \Cref{tab:ablation}, we used YAGO3-10 as the benchmark dataset and the AnyBURL 50s ruleset (107k rules).
-The “tables for each relation” optimization specifically reduced the execution time for long rules.
-For rules with one body atom, the execution time was reduced by 12\%, for two body atoms by 25\% and for three body atoms by 74\%.
-This is caused by the increased use of the indexes during execution for longer rules.
+For the ablation study in \Cref{tab:ablation}, we use YAGO3-10 as the benchmark dataset and the AnyBURL 50s ruleset (107k rules).
+
+The “tables for each relation” optimization described in \ref{database-structure} specifically reduces the execution time for long rules.
+We tested this, by measuring the average execution time of different rule lengths in the version where the “table for each relation” optimization is enabled versus the version where the “table for each relation” optimization is disabled.
+In both runs, the rule pre-computing was disabled.
+In this experiment, the execution time for rules with one body atom is reduced by 12\%, for two body atoms by 25\% and for three body atoms by 74\% when the “table for each relation” optimization is enabled.
+This might be caused by the increased use of the indexes during execution for longer rules.
 
@@
 	ylabel=\#appearances * execution time (ms),
-        xlabel=Percentile,
+        xlabel=Percentile of rules,
 	enlargelimits=0.05,
@@
 
-The quantile analysis in \Cref{fig:quantile-analysis} is based on the number of appearances multiplied with the average execution of the rules after using the validation set as target triples as described in \ref{pre-computation}.
+The quantile analysis in \Cref{fig:quantile-analysis} measures the total cost per rule after using the validation set as target triples, as described in \ref{pre-computation}.
+The total cost per rule is calculated by multiplying the number of appearances with the average execution time.
+Here, the number of appearances is a count for how often a rule appears as a result of the queries for all target triples.
 It indicates that only few rules incur a major share of the cost.
@@
 
-We demonstrated an effective approach for finding all rules that imply a certain target fact given a knowledge graph and a set of previously learned rules.
-Our experiments specifically demonstrated the effect of filtering, indexing and pre-computing methods.
+We have designed an effective approach for finding all rules that entail a certain target fact given a knowledge graph and a set of previously learned rules.
+Our experiments specifically demonstrate the effect of indexing, filtering and pre-computing methods.
 Potential next steps include a further analysis of our approach on various datasets, a comparison of different database technologies (particularly triplestores), exploring the use of multithreading and creating a dedicated solution only using the main memory.

### 2022-12-17 Paper/Overleaf/paper.tex
Paper - Final Version #2

@@
 %%% Das optionale Argument (sofern angegeben) wird für die Kopfzeile verwendet.
-\title[Which Rules Entail this Fact? An Efficient Approach Using RDBMSs]{Which Rules Entail this Fact? An Efficient Approach Using RDBMSs}
+\title[Which Rules Entail this Fact? An Efficient Approach Using RDBMSs]{Which Rules Entail this Fact? \newline An Efficient Approach Using RDBMSs}
 %%% \subtitle{From Facts to Rules using Relational Databases} % if needed
@@
 
-The given KG is stored in one table of the relational database (\textit{kg\_table} with columns “sub” for subjects, “rel” for relations and “obj” for objects).
+The given KG is stored in one table of the relational database (\textit{kg\_table} with columns \textit{sub} for subjects, \textit{rel} for relations and \textit{obj} for objects).
 The basic idea behind our approach is creating SQL queries which check whether a certain rule entails a certain target fact.
@@
 The easiest way is storing each rule under its head as key in a hash map.
-The head stays unchanged, but we use “X” and “Y” as variable descriptors.
+The head stays unchanged, but we use “\textit{X}” and “\textit{Y}” as variable descriptors.
 
-Let (s, r, o) be a target triple with the entities s and o and the relation r.
-Then we only use the rules stored under the keys \textit{(X, r, Y)}, \textit{(s, r, X)}, \textit{(X, r, o)} or \textit{(s, r, o)} for the query.
-Additionally, when it is true that s=o for the target, we also use the key \textit{(X, r, X)}.
-Instead of looping through n rules naively in O(n), this allows for initial filtering of the rules in O(1).
+Let \textit{(s, r, o)} be a target triple with the entities \textit{s} and \textit{o} and the relation \textit{r}.
+Then we only use the rules stored under the keys \textit{(X, r, Y)}, \textit{(s, r, X)}, \textit{(X, r, o)} or \textit{(s, r, o)} for the query of the target triple.
+Additionally, when it is true that \(s=o\), we also use the key \textit{(X, r, X)}.
+Instead of looping through n rules naively in O(n), this allows for pre-filtering of the rules in O(1).
 
@@
 If the subject or the object is a variable and the other one a constant, then the table only contains the column for the variable.
-For the implementation of the tables, we use materialized views to \(SELECT\) from the KG with all the conditions we already know from the rule.
+For the implementation, we use indexed materialized views to \(SELECT\) from the KG with all the conditions we already know from the rule.
 Consequently, the \(CREATE\) statement for pre-computing rule (\ref{eq:rule_2}) and the query for the target fact from the preliminaries would be:
@@
 \end{tabular}
-\caption{Composition of the rules (\# of rules | avg. share by \# of body atoms)}
+\caption{Composition of the rules (\# of rules | avg. share of rules by \# of body atoms)}
 \label{tab:rules}

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java
make Cyclic rules measureable

@@ -57,6 +57,7 @@ public class RandomRules {
     }
 
     public void importRules() {
+        boolean acyclic_continuer = false;
         Key3Int key3Int;
         int oneAtomCounter = 0;
         int twoAtomCounter = 0;
@@ -164,7 +165,21 @@ public class RandomRules {
                 } else {
                     continue;
                 }
-
+                acyclic_continuer = false;
+                if(Config.getStringValue("ONLY_ACYCLIC").equals("YES")){
+                    if(rule.getBody().size() <= 1){
+                        continue;
+                    }
+                    if(rule.getHead().getSubject() == rule.getBody().get(0).getSubject() && rule.getHead().getObject() == rule.getBody().get(rule.getBody().size()-1).getObject()){
+                        acyclic_continuer = true;
+                    }
+                    if(rule.getHead().getObject() == rule.getBody().get(0).getSubject() && rule.getHead().getSubject() == rule.getBody().get(rule.getBody().size()-1).getObject()){
+                        acyclic_continuer = true;
+                    }
+                    if(!acyclic_continuer){
+                        continue;
+                    }
+                }
 
                 if (Config.getStringValue("FILTER_SIMPLE_RULES").equals("YES")) {
                     continuer = false;
@@ -277,7 +292,9 @@ public class RandomRules {
                 System.out.println("3 Atom rules " + threeAtomCounter);
                 System.out.println("4 Atom rules " + fourAtomCounter);
                 System.out.println("5 Atom rules " + fiveAtomCounter);
+                System.out.println("Total rule Count " + counter);
                 System.out.println("Rule Import finished");
+                //System.exit(0);
             //System.exit(0);
             reader.close();
             if (Config.getStringValue("TESTRULES_METHOD").equals("testRulesSimpleViews") && Config.getStringValue("REFILL_TABLES").equals("YES")) {
@@ -417,7 +434,7 @@ public class RandomRules {
         if (ruleSet != null) {
             filteredRules.addAll(ruleSet);
         }
-        //ilteredRules.forEach(e -> System.out.println(e.getId() + " : " + e));
+        //filteredRules.forEach(e -> System.out.println(e.getId() + " : " + e));
         ArrayList<Integer> resultList;
         switch (Config.getStringValue("TESTRULES_METHOD")) {
             case "testRulesUnionAllShorterSelect":

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java
update config.properties

@@ -167,13 +167,7 @@ public class RandomRules {
                 }
                 acyclic_continuer = false;
                 if(Config.getStringValue("ONLY_ACYCLIC").equals("YES")){
-                    if(rule.getBody().size() <= 1){
-                        continue;
-                    }
-                    if(rule.getHead().getSubject() == rule.getBody().get(0).getSubject() && rule.getHead().getObject() == rule.getBody().get(rule.getBody().size()-1).getObject()){
-                        acyclic_continuer = true;
-                    }
-                    if(rule.getHead().getObject() == rule.getBody().get(0).getSubject() && rule.getHead().getSubject() == rule.getBody().get(rule.getBody().size()-1).getObject()){
+                    if(rule.getHead().getSubject() == -24 && rule.getHead().getObject() == -25){
                         acyclic_continuer = true;
                     }
                     if(!acyclic_continuer){

### 2022-12-19 Code/sql-implementations/src/app/Main.java
update config.properties

@@ -20,6 +20,8 @@ public class Main {
   private static ConnectDB connectDB = new ConnectDB(Config.getStringValue("CLASSNAME"), Config.getStringValue("URL"), Config.getStringValue("USER"), Config.getStringValue("PASSWORD"));
 
   public static void main(String[] args) {
+    System.out.println("Hello World");
+    System.exit(0);
     initalize();
     DbFill dbFill = new DbFill();
     ArrayList<Integer> arrayList = new ArrayList<>();

### 2022-12-19 Code/sql-implementations/src/app/Main.java
update JAR

@@ -20,8 +20,6 @@ public class Main {
   private static ConnectDB connectDB = new ConnectDB(Config.getStringValue("CLASSNAME"), Config.getStringValue("URL"), Config.getStringValue("USER"), Config.getStringValue("PASSWORD"));
 
   public static void main(String[] args) {
-    System.out.println("Hello World");
-    System.exit(0);
     initalize();
     DbFill dbFill = new DbFill();
     ArrayList<Integer> arrayList = new ArrayList<>();

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java
update JAR

@@ -761,7 +761,7 @@ public class RandomRules {
         ArrayList<Double> learnedTimelist;
         ArrayList<Double> timeListAll = new ArrayList<>();
         int i = 0;
-        int threshhold = (int) (ruleCount * 0.01);
+        int threshhold = (int) (ruleCount * 0.10);
         //System.out.println("Rule Count: " + ruleCount);
         //System.out.println("Threshhold " + threshhold);
         //System.exit(0);

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java
update JAR

@@ -170,7 +170,7 @@ public class RandomRules {
                     if(rule.getHead().getSubject() == -24 && rule.getHead().getObject() == -25){
                         acyclic_continuer = true;
                     }
-                    if(!acyclic_continuer){
+                    if(acyclic_continuer){
                         continue;
                     }
                 }

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java
update JAR

@@ -170,7 +170,10 @@ public class RandomRules {
                     if(rule.getHead().getSubject() == -24 && rule.getHead().getObject() == -25){
                         acyclic_continuer = true;
                     }
-                    if(acyclic_continuer){
+                    if(rule.getBody().size() != 3){
+                        continue;
+                    }
+                    if(!acyclic_continuer){
                         continue;
                     }
                 }

### 2022-12-19 Code/sql-implementations/src/database/DBFuncs.java
update JAR

@@ -1112,8 +1112,8 @@ public class DBFuncs {
                 preSelect.append(select);
                 preSelect.append(where);
                 sql.append(preSelect);
-                //System.out.println(rule);
-                //System.out.println(sql);
+                System.out.println(rule);
+                System.out.println(sql);
                 stmt.addBatch(sql.toString());
                 sql = new StringBuffer("CREATE UNIQUE INDEX x" + rule.getId() + "Index ON x" + rule.getId());
                 //System.out.println("Subset, Objectset: " + subSet + ", " + objSet);
@@ -1135,7 +1135,7 @@ public class DBFuncs {
                 sql = new StringBuffer("ALTER MATERIALIZED VIEW x" + rule.getId() + " CLUSTER ON x" + rule.getId() + "Index;");
                 stmt.addBatch(sql.toString());
                 count++;
-                if (count % 5 == 0 || count == rules.size()) {
+                if (count % 1 == 0 || count == rules.size()) {
                     stmt.executeBatch();
                     stmt.clearBatch();
                     elapsedTime = System.nanoTime() - startTime;

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java

@@ -166,14 +166,14 @@ public class RandomRules {
                     continue;
                 }
                 acyclic_continuer = false;
-                if(Config.getStringValue("ONLY_ACYCLIC").equals("YES")){
-                    if(rule.getHead().getSubject() == -24 && rule.getHead().getObject() == -25){
+                if (Config.getStringValue("ONLY_ACYCLIC").equals("YES")) {
+                    if (rule.getHead().getSubject() == -24 && rule.getHead().getObject() == -25) {
                         acyclic_continuer = true;
                     }
-                    if(rule.getBody().size() != 3){
+                    if (rule.getBody().size() != 3) {
                         continue;
                     }
-                    if(!acyclic_continuer){
+                    if (!acyclic_continuer) {
                         continue;
                     }
                 }
@@ -268,30 +268,35 @@ public class RandomRules {
 
                 rule.setId(counter++);
                 //System.out.println(counter);
-                if(rule.getBody().size() == 1) {
+                if (rule.getBody().size() == 1) {
                     oneAtomCounter++;
-                };
-                if(rule.getBody().size() == 2) {
+                }
+                ;
+                if (rule.getBody().size() == 2) {
                     twoAtomCounter++;
-                };
-                if(rule.getBody().size() == 3) {
+                }
+                ;
+                if (rule.getBody().size() == 3) {
                     threeAtomCounter++;
-                };
-                if(rule.getBody().size() == 4) {
+                }
+                ;
+                if (rule.getBody().size() == 4) {
                     fourAtomCounter++;
-                };
-                if(rule.getBody().size() == 5) {
+                }
+                ;
+                if (rule.getBody().size() == 5) {
                     fiveAtomCounter++;
-                };
+                }
+                ;
             }
-                System.out.println("1 Atom rules " + oneAtomCounter);
-                System.out.println("2 Atom rules " + twoAtomCounter);
-                System.out.println("3 Atom rules " + threeAtomCounter);
-                System.out.println("4 Atom rules " + fourAtomCounter);
-                System.out.println("5 Atom rules " + fiveAtomCounter);
-                System.out.println("Total rule Count " + counter);
[…]

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java
update Config

@@ -767,7 +767,8 @@ public class RandomRules {
         ArrayList<Double> learnedTimelist;
         ArrayList<Double> timeListAll = new ArrayList<>();
         int i = 0;
-        int threshhold = (int) (ruleCount * 0.10);
+        //int threshhold = (int) (ruleCount * 0.10);
+        int threshhold = 50;
         //System.out.println("Rule Count: " + ruleCount);
         //System.out.println("Threshhold " + threshhold);
         //System.exit(0);

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java
update JAR

@@ -296,7 +296,7 @@ public class RandomRules {
             System.out.println("5 Atom rules " + fiveAtomCounter);
             System.out.println("Total rule Count " + counter);
             System.out.println("Rule Import finished");
-            //System.exit(0);
+            // System.exit(0);
             //System.exit(0);
             reader.close();
             if (Config.getStringValue("TESTRULES_METHOD").equals("testRulesSimpleViews") && Config.getStringValue("REFILL_TABLES").equals("YES")) {

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java
update JAR

@@ -296,7 +296,7 @@ public class RandomRules {
             System.out.println("5 Atom rules " + fiveAtomCounter);
             System.out.println("Total rule Count " + counter);
             System.out.println("Rule Import finished");
-            // System.exit(0);
+            System.exit(0);
             //System.exit(0);
             reader.close();
             if (Config.getStringValue("TESTRULES_METHOD").equals("testRulesSimpleViews") && Config.getStringValue("REFILL_TABLES").equals("YES")) {

### 2022-12-19 Code/sql-implementations/src/app/Main.java
update JAR

@@ -29,6 +29,7 @@ public class Main {
     DBFuncs.viewsForPredicate(arrayList);*/
     if(Config.getStringValue("REFILL_TABLES").equals("YES")){
       dbFill.fillVocabulary();
+      System.out.println("Filled Vocabulary");
       dbFill.fillKnowledgegraph();
       System.out.println("Knowledgegraph filled");
     }else {

### 2022-12-19 Code/sql-implementations/src/utils/DbFill.java

@@ -192,6 +192,7 @@ public class DbFill {
         break;
       default:
     }
+    System.out.println("Before Fill Indizes");
     fillIndizes();
     System.out.println("Vocabulary filled");
   }

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java

@@ -296,7 +296,7 @@ public class RandomRules {
             System.out.println("5 Atom rules " + fiveAtomCounter);
             System.out.println("Total rule Count " + counter);
             System.out.println("Rule Import finished");
-            System.exit(0);
+            //System.exit(0);
             //System.exit(0);
             reader.close();
             if (Config.getStringValue("TESTRULES_METHOD").equals("testRulesSimpleViews") && Config.getStringValue("REFILL_TABLES").equals("YES")) {
@@ -791,14 +791,15 @@ public class RandomRules {
         }
         //quantilCalcSum(timeListAll);
         //System.exit(0);
-        DBFuncs.viewsForQuantiles(ruleList);
         rulePreSave(ruleList);
+        DBFuncs.viewsForQuantiles(ruleList);
         System.out.println("Quantil Calc old Rule learning");
         learnedTimelist = getLearnedRuleTime(preProcessedRules, this.validationTriples);
         System.out.println("Quantil Calc new Rule Learning for all Rules");
         printPriciestRuleTimes(timeList);
         //quantilCalcSum(learnedTimelist);
         printPriciestRuleTimes(learnedTimelist);
+        System.exit(0);
         try {
             Thread.sleep(30);
         } catch (InterruptedException e) {

### 2022-12-19 Code/sql-implementations/src/database/DBFuncs.java
update JAR

@@ -1677,6 +1677,7 @@ public class DBFuncs {
             }
             stmt.executeBatch();
             con.commit();
+            System.out.println("Commit dropping all Materialized views");
             count = 0;
             startTime = System.nanoTime();
             for (Integer integer : predicateList) {

### 2022-12-19 Code/sql-implementations/src/utils/RandomRules.java
update JAR

@@ -750,12 +750,14 @@ public class RandomRules {
             filteredRules.addAll(ruleSet);
         }
         for (Rule rule : filteredRules) {
+            if(rule.getBody().size() == 3 && rule.getHead().getSubject() == -24 && rule.getHead().getObject() == -25){
                 System.out.println(ruleLearnCount++);
                 if (ruleTimeHashMap.containsKey(rule.getId())) {
                     ruleTimeHashMap.get(rule.getId()).addTime(DBFuncs.timePerRule(rule, triple));
                 } else {
                     ruleTimeHashMap.put(rule.getId(), new RuleTime(DBFuncs.timePerRule(rule, triple), rule));
                 }
+            }
         }
     }
 

### 2023-02-14 Paper/Overleaf/paper.tex
Updated paper based on feedback from the PC committee

@@
 
-\pgfplotsset{width=12.5cm, height=5.3cm}
+\pgfplotsset{width=12.5cm, height=4.4cm}
 
@@
 %%% Das optionale Argument (sofern angegeben) wird für die Kopfzeile verwendet.
-\title[Which Rules Entail this Fact? An Efficient Approach Using RDBMSs]{Which Rules Entail this Fact? \newline An Efficient Approach Using RDBMSs}
+\title[Which Rules Entail this Fact? - An Efficient Approach Using RDBMSs]{Which Rules Entail this Fact? \newline An Efficient Approach Using RDBMSs}
 %%% \subtitle{From Facts to Rules using Relational Databases} % if needed
@@
 \startpage{1} % Beginn der Seitenzählung für diesen Beitrag / Start page
-\editor{B. König-Ries et al.} % Names of Editors
-\booktitle{Datenbanksysteme für Business, Technologie und Web (BTW 2023)} % Name of book title
+\editor{B. K{"o}nig-Ries et al.} % Names of Editors
+\booktitle{Datenbanksysteme f{"u}r Business, Technologie und Web (BTW 2023)} % Name of book title
 \yearofpublication{2023}
@@
 \begin{abstract}
- In this paper, we focus on the problem of identifying all rules that entail a certain target fact given a knowledge graph and a set of previously learned rules.
-This problem is relevant in the context of link prediction and explainability.
-We propose an efficient approach using relational database technology including indexing, filtering and pre-computing methods.
+Knowledge graphs (KGs) are used to store information about relationships between real-world entities in various fields.
+Learned rules over KGs describe patterns of KGs and allow for knowledge inference.
+In this paper, we focus on the problem of identifying all rules that entail a certain target fact given a KG and a set of previously learned rules.
+This can enable link prediction as well as help explain connections between rules and (potential) facts.
+Solving this problem time-efficiently for large rulesets and KGs is a challenge.
+To tackle this challenge, we propose an approach relying solely on RDBMSs including indexing, filtering and pre-computing methods.
 Our experiments demonstrate the efficiency of our approach and the effect of various optimizations on different datasets like YAGO3-10, WN18RR and FB15k-237 using rules learned by the bottom up rule learner AnyBURL.
@@
 \section{Introduction}
-Knowledge graphs (KGs) are used in different fields from biomedical applications \cite{OpenBioLink} over applications in the context of social networks \cite{SocialNetworks} to semantic search applications \cite{SemanticSearch}.
-Learned rules over the KG describe patterns of the KG and allow for knowledge inference.
-This can be for example used to predict missing or additional information also known as the link prediction problem.
-
-Identifying which previously learned rules of a KG entail certain target facts is an important problem for two main reasons.
-Firstly, to determine the confidence of an (unknown) fact, one must know from which rules it can be derived.
+Many practically relevant large KGs are incomplete.
+Therefore, the prediction of missing or additional information, also known as link prediction, is highly relevant in contexts such as biomedicine \cite{OpenBioLink} or social networks \cite{SocialNetworks}.
+Rules describing patterns of KGs, which are learned by rule learning tools, can enable link prediction.
+The rule \((X, citizenOf, germany) \leftarrow (X, bornIn, mannheim)\) describes for example the implication that someone (\(X\)) born in Mannheim is a citizen of Germany.
+To determine the confidence of a potential new fact (link), one must know from which rules it can be derived.
 The confidence of a fact refers to the probability that a particular fact within the graph is correct.
+In our example: If we know that \(X\) was born in Mannheim it can help us determine the confidence of the fact that \(X\) is a citizen of Germany.
 Knowing the confidence is needed for rule-based link prediction models \cite{AnyBURL19}.
-Moreover, the confidence is used for building ensembles with embedding models \cite{RuleEmbeddingCombination1} and for improving embedding models \cite{RuleEmbeddingCombination2}.
-Secondly, it is generally helpful to be able to explain the connections between rules and (potential) facts.
-There is a significant need for a time-efficient approach to solve this problem.
-
-In this paper, we create and analyze an effective approach to this problem, using relational database technology.
+Moreover, the confidence is used for building ensembles with link prediction embedding models \cite{RuleEmbeddingCombination1} and for improving link prediction embedding models \cite{RuleEmbeddingCombination2}.
+Therefore, it is important to identify all previously learned rules of a KG entailing certain target facts.
+Moreover, it is generally helpful to be able to explain the connections between rules and (potential) facts.
+To illustrate the performance of traditional RDBMSs in solving the problem in case of large rulesets and KGs, we create and analyze an efficient approach to this problem using PostgreSQL.
+As further discussed in the preliminaries, we are interested in the non-recursive application of rules and in finding all applicable rules for a given target fact.
+Therefore, our approach is limited to binary relations and conjunctive queries, in contrast to employing deductive databases.
 Our approach uses indexing, filtering and pre-computing methods tailored to the natural structure of a KG and its respective rules.
-To illustrate the performance of our approach, we ran several experiments on different KGs (YAGO3-10 \cite{YAGO3}, WN18RR \cite{WN18RR} and FB15k-237 \cite{FB15k-237}).
-Moreover, we show how different database setups impact the performance for different rule lengths.
+We ran several experiments testing our approach on different KGs (YAGO3-10 \cite{YAGO3}, WN18RR \cite{WN18RR} and FB15k-237 \cite{FB15k-237}).
+Additionally, we show how different database setups impact the performance for different rule lengths.
 For the rule learning, we use AnyBURL, a fast bottom up rule learner for KGs \cite{AnyBURL19}.
@@
 A KG is a set of (subject, relation, object)-triples also called facts.
-There is a set of entities present as subjects and objects as well as a set of relations in a KG.
-Here is an example KG with the entities \(peter\), \(anna\) and \(germany\) and the relations \(livesIn\) and \(marriedTo\).
+There is a set of entities present as subjects and objects, as well as a set of relations in a KG.
+Here is an example KG with the entities \(peter\), \(anna\) and \(germany\) and the relations \(marriedTo\) and \(bornIn\).
 
@@
 To improve this basic idea, we employ several optimizations listed below.
+Every rule has a specific rule ID which is returned if the rule entails the target fact.
 This would be the query for rule (\ref{eq:rule_2}) and the target fact given in the preliminaries:
 
-\textit{\textbf{SELECT} rule\_1 \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} t0.sub = anna \textbf{AND} t0.rel = marriedTo \textbf{AND} t0.obj = t1.sub \textbf{AND} t1.rel = bornIn \textbf{AND} t1.obj = germany \textbf{LIMIT} 1;}
+\textit{\textbf{SELECT} rule\_id\_2 \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} t0.sub = anna \textbf{AND} t0.rel = marriedTo \textbf{AND} t0.obj = t1.sub \textbf{AND} t1.rel = bornIn \textbf{AND} t1.obj = germany \textbf{LIMIT} 1;}
 
@@
 Alternative to having one big table for all facts, our approach uses one table for each relation in the KG, with two columns for the subjects and the objects.
-This improves the performance by reducing the size of the B-trees of the indexes described in the following paragraph.
 
@@
 \label{pre-computation}
-As we will illustrate in our experiments, certain rules result in way more cost in terms of execution time than others.
+As we will illustrate in our experiments (\ref{quintile-analysis}), certain rules result in way more cost in terms of execution time than others.
 To address that, we pre-compute a portion of those rules.
 To identify the most expensive rules, we gather an independent set of n target facts from the KG and run their queries with the \textit{EXPLAIN ANALYZE} command to get the execution time for the individual rule sub-queries.
-Afterwards, we rank all rules which appear in the results by their average execution time multiplied with their number of appearances in the results.
+Afterward, we rank all rules which appear in the results by their average execution time multiplied with their number of appearances in the results.
 
@@
 
-\textit{\textbf{CREATE MATERIALIZED VIEW} view\_rule\_1 \textbf{SELECT} t0.sub \textbf{AS} sub \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} t0.rel = marriedTo \textbf{AND} t0.obj = t1.sub \textbf{AND} t1.rel = bornIn \textbf{AND} t1.obj = germany};
+\textit{\textbf{CREATE MATERIALIZED VIEW} view\_rule\_2 \textbf{SELECT} t0.sub \textbf{AS} sub \textbf{FROM} kg\_table t0,  kg\_table t1 \textbf{WHERE} t0.rel = marriedTo \textbf{AND} t0.obj = t1.sub \textbf{AND} t1.rel = bornIn \textbf{AND} t1.obj = germany};
 
-\textit{\textbf{SELECT} rule\_1 \textbf{FROM} view\_rule\_1  \textbf{WHERE} sub = anna \textbf{LIMIT} 1};
+\textit{\textbf{SELECT} rule\_id\_2 \textbf{FROM} view\_rule\_2  \textbf{WHERE} sub = anna \textbf{LIMIT} 1};
 
 We exclude rules with one body atom, as pre-computing does not improve their performance.
-Moreover, we exclude rules with more than two body atoms, as those rules incur a pre-computing time two orders of magnitude higher than rules with two body atoms, while they only make up a fraction of the most expensive and frequent rules.
+Beyond that, we limit the pre-computation to rules with two body atoms for illustration purposes, as rules with more body atoms incur a pre-computation time two orders of magnitude higher than rules with two body atoms.
 
@@
 The goal of our experiments is to benchmark the performance of our approach on different datasets and rulesets, as well as to measure the effects of different optimizations.
-Beyond that, we analyze the execution time for individual rules.
+Furthermore, we analyze the execution time for individual rules.
 
@@
 We only extended the limit of body atoms from one to three for acyclic rules to match the cyclic rules, as we do not intend to discriminate between them.
+This means the rulesets only include rules with up to three body atoms.
+The relations in the rule bodies are always extensional (defined by facts), never intensional (defined by other rules).
 
@@
 \centering
-\begin{tabular}{l|lll|lll}
+\begin{tabular}{l|lll}
 \toprule
- & 10s & 50s  & 100s & \#entities & \#relations & \#facts of KG\\
+ & AnyBURL - 10s & AnyBURL - 50s  & AnyBURL - 100s\\
 \midrule
-YAGO3-10 & 4 ms & 10 ms & 14 ms & 123,182 & 37 & 1,079,040\\
-WN18RR & 35 ms & 65 ms & 77 ms & 40,943 & 11 & 86,835\\
-FB15k-237 & 5 ms & 26 ms & 43 ms & 14,505 & 237 & 272,115\\
+YAGO3-10 & 4.2 & 10.4 & 14.3\\
+WN18RR & 34.8 & 65.0 & 77.3\\
+FB15k-237 & 5.4 & 26.2 & 42.9\\
 \bottomrule
 \end{tabular}
-\caption{Performance results (avg. execution time per target fact | dataset information)}
+\caption{Performance results (avg. execution time per target fact in ms)}
 \label{tab:overall}
@@
 \toprule
- & 10s & 50s & 100s & 1 body atom & 2 body atoms & 3 body atoms\\
+ & \#entities & \#relations & \#facts of KG & 10s & 50s & 100s\\
 \midrule
-YAGO3-10 & 27k & 107k & 165k & 82.8\% & 11.2\% & 6.0\%\\
-WN18RR & 6k & 20k & 30k & 33.0\% & 23.9\% & 43.1\%\\
-FB15k-237 & 33k & 135k & 252k & 53.9\% & 26.2\% & 19.9\%\\
+YAGO3-10 & 123,182 & 37 & 1,079,040 & 27k & 107k & 165k\\
+WN18RR & 40,943 & 11 & 86,835 & 6k & 20k & 30k\\
+FB15k-237 & 14,505 & 237 & 272,115 & 33k & 135k & 252k\\
 \bottomrule
 \end{tabular}
-\caption{Composition of the rules (\# of rules | avg. share of rules by \# of body atoms)}
+\caption{Dataset properties \& \#rules per ruleset learned by AnyBURL}
 \label{tab:rules}
@@
 For every dataset, we learned rules for 10s, 50s, and 100s (\Cref{tab:rules}).
-We use the training sets as the KGs and the test sets as the target triples (3k-20k triples).
-The validation sets (3k-20k triples) are used as target triples to create the rule rankings to pre-compute the top 1\% most expensive rules.
+We used the training sets as the KGs and the test sets as the target triples (3k-20k triples).
+The validation sets (3k-20k triples) were used as target triples to create the rule rankings to pre-compute the top 1\% most expensive rules with two body atoms.
 
@@
 \toprule
-Experiment & Avg. execution time \\
+Experiment & Avg. execution time (in ms) \\
 \midrule
-All optimizations enabled & 10 ms\\
+All optimizations enabled & 10.4\\
 \midrule
-Advanced rule pre-filtering disabled & 55 ms\\
-Tables for each relation disabled & 133 ms\\
-Indexing disabled & 1720 ms\\
-Pre-computing of expensive rules disabled & 16 ms\\
+Advanced rule pre-filtering disabled & 55.0\\
+Tables for each relation disabled & 132.9\\
+Indexing disabled & 1720.2\\
+Pre-computing of expensive rules disabled & 14.1\\
 
@@
 
-For the ablation study in \Cref{tab:ablation}, we use YAGO3-10 as the benchmark dataset and the AnyBURL 50s ruleset (107k rules).
-
+For the ablation study in \Cref{tab:ablation}, we used YAGO3-10 as the benchmark dataset and the AnyBURL 50s ruleset (107k rules).
+We didn't analyze versions with multiple optimizations disabled due to very high execution times.
 The “tables for each relation” optimization described in \ref{database-structure} specifically reduces the execution time for long rules.
 We tested this, by measuring the average execution time of different rule lengths in the version where the “table for each relation” optimization is enabled versus the version where the “table for each relation” optimization is disabled.
-In both runs, the rule pre-computing was disabled.
-In this experiment, the execution time for rules with one body atom is reduced by 12\%, for two body atoms by 25\% and for three body atoms by 74\% when the “table for each relation” optimization is enabled.
+In both runs, the rule pre-computation was disabled.
+In this experiment, the execution time for rules with one body atom was reduced by 12\%, for two body atoms by 25\% and for three body atoms by 74\% when the “table for each relation” optimization is enabled.
 This might be caused by the increased use of the indexes during execution for longer rules.
@@
 \subsection{Rule Quantile Performance Analysis} 
+\label{quintile-analysis}
 
@@
 		/pgf/number format/1000 sep=},
-	ylabel=\#appearances * execution time (ms),
+	ylabel=total cost per rule (ms),
         xlabel=Percentile of rules,
@@
 
-The quantile analysis in \Cref{fig:quantile-analysis} measures the total cost per rule after using the validation set as target triples, as described in \ref{pre-computation}.
+The quantile analysis in \Cref{fig:quantile-analysis} is based on the total cost per rule after using the validation set as target triples, as described in \ref{pre-computation}.
 The total cost per rule is calculated by multiplying the number of appearances with the average execution time.
@@
 It indicates that only few rules incur a major share of the cost.
-This ranking is used to determine the top x\% for the “pre-computing of expensive rules” optimization.
-In the given example (YAGO3-10 \& 107k rules), we achieved an 98.8\% execution time reduction for the pre-computed top 1\% rules.
-The pre-compution took five minutes.
-This reduced our average execution time from 16 ms to 10 ms, as illustrated in \Cref{tab:ablation}.
+The most expensive rules are predominantly rules with two variables in the rule head and two or three body atoms.
+Excluding those rules (978 rules) reduced the average execution time for YAGO3-10 \& the AnyBURL 50s ruleset (107k rules) to 0.14 ms per query using all optimizations.
+In the given example (YAGO3-10 \& 107k rules), we achieved an 98.8\% execution time reduction for the pre-computed rules.
+The pre-computation took 15 minutes.
+This reduced our average execution time from 14.1 ms to 10.4 ms, as illustrated in \Cref{tab:ablation}.
 
@@
 
-We have designed an effective approach for finding all rules that entail a certain target fact given a knowledge graph and a set of previously learned rules.
+We have designed an efficient approach for finding all rules that entail a certain target fact given a knowledge graph and a set of previously learned rules.
 Our experiments specifically demonstrate the effect of indexing, filtering and pre-computing methods.
-Potential next steps include a further analysis of our approach on various datasets, a comparison of different database technologies (particularly triplestores), exploring the use of multithreading and creating a dedicated solution only using the main memory.
+Potential next steps include a further analysis of our approach on various datasets, an empirical comparison of different database technologies (particularly triplestores and deductive DBMS), exploring the use of multithreading, investigating the use of multidimensional indexes and creating a dedicated solution only using the main memory.
 
